### 2012-10-14 supp/sim_table_exp.tex
rough simulation results plus

\relax 
\providecommand\HyperFirstAtBeginDocument{\AtBeginDocument}
\HyperFirstAtBeginDocument{\ifx\hyper@anchor\@undefined
\global\let\oldcontentsline\contentsline
\gdef\contentsline#1#2#3#4{\oldcontentsline{#1}{#2}{#3}}
\global\let\oldnewlabel\newlabel
\gdef\newlabel#1#2{\newlabelxx{#1}#2}
\gdef\newlabelxx#1#2#3#4#5#6{\oldnewlabel{#1}{{#2}{#3}}}
\AtEndDocument{\ifx\hyper@anchor\@undefined
\let\contentsline\oldcontentsline
\let\newlabel\oldnewlabel
\fi}
\fi}
\global\let\hyper@last\relax 
\gdef\HyperFirstAtBeginDocument#1{#1}
\providecommand*\HyPL@Entry[1]{}
\HyPL@Entry{0<</S/D>>}
\@writefile{lot}{\contentsline {table}{\numberline {1}{\ignorespaces Simulation results from 27 cases of trend dynamics with 1000 Monte Carlo replicates for each.}}{1}{table.1}}
\newlabel{tab:simexp}{{1}{1}{Simulation results from 27 cases of trend dynamics with 1000 Monte Carlo replicates for each}{table.1}{}}