### 2013-02-26 Dail_Mad_ext.tex
Several minor edits2

\documentclass[12pt]{article}

\usepackage[a4paper,hmargin={2.4cm,2.4cm},vmargin={2.4cm,2.4cm}]{geometry}
\usepackage[mathlines,displaymath]{lineno}
\usepackage{setspace}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{bm}
\usepackage{graphicx}
\usepackage[round,colon,authoryear]{natbib}
\usepackage{color, soul}
\usepackage{fancyvrb}
\usepackage{verbatim}
\usepackage[pdftex,colorlinks=true,citecolor=black,linkcolor=black,urlcolor=black,pdftex,pdfstartview=FitH,bookmarks=true]{hyperref}
\usepackage{rotating}

\bibliographystyle{ecology}

\begin{document}

\title{Improved state-space models for inference about spatial
and temporal variation in abundance from count data}

\author{Jeffrey A. Hostetler\\ %\footnote{Contact:
%     \url{HostetlerJ@si.edu}, +001-202-633-4205}\\
%        \small Migratory Bird Center \\
        \small Smithsonian Conservation Biology Institute \\
        \small National Zoological Park, MRC 5503 \\
        \small Washington, DC 20013-7012\\
   \and Richard B. Chandler \footnote{Contact:
     \url{RChandler@usgs.gov}, +001-301-497-5856} \\
        \small USGS Patuxent Wildlife Research Center \\
        \small Laurel, MD 20708}
\maketitle

\abstract{
Models of population dynamics play a central role in theoretical and
applied ecology where they are used for purposes such as testing 
hypotheses about density dependence and predicting species' responses 
to future environmental change or conservation actions. Failure to
account for observation error in such models can result in 
bias, and thus ecologists have increasingly relied on state-space
models to directly model both the observation error and the ecological
process of interest.
Conventional state-space models, however, have
three important limitations: (1) the parameters are not identifiable
in many common situations, (2) they do not admit spatial variation in
population dynamics, and (3) there is no clear interpretation of the
observation error. We demonstrate how each of these problems can be
resolved using a class of hierarchical models proposed by
\citet[Biometrics]{dail_madsen:2011} for spatially-replicated
time-series data that attributes observation error to imperfect detection.
We expand this class of models to accommodate classical growth
models (e.g. exponential and Ricker-logistic), zero-inflation,
and random effects such as observer-specific
detection probabilities. We also present methods for forecasting
population size under future environmental conditions. Implementation
of these ideas is possible using either frequentist or Bayesian
methods, and code to fit these models using the \textbf{R} package
\texttt{unmarked} and \textbf{JAGS} is also provided.
A simulation study indicated that bias was negligible and coverage
nominal for the proposed model extensions. An analysis of data from
the North American Breeding Bird Survey highlighted how these methods
can be readily applied to existing data, but it also suggested the
that precision will be low when direct information about detection
probability (such as is collected using distance sampling or
replicated counts) is unavailable.
}

\vspace{0.5cm}

\textbf{Key words}: abundance, Dail and Madsen model, density-dependence,
Gompertz-logistic model, immigration, open population point count models,
random observer effects, range, Ricker-logistic model, zero-inflated

\newpage

Theoretical ecology requires models of population dynamics for testing
hypotheses regarding spatial and temporal variation in
abundance. For example, much theoretical work has focused on
understanding the importance and existence of phenomenon such as
density-dependent population regulation, population cycling, and spatial synchrony
\citep{may:1975,royama:1977,turchin:1990,dennis_taper:1994,bjornstad_etal:1999},
and population models are required to evaluate associated hypotheses.
In applied contexts, population models are used for estimating extinction probabilities
\citep{schoener_spiller:1992,nadeem_lele:2011,hostetler_etal:2012} and for predicting the
effects of future environmental conditions or conservation actions on
population size \citep{jamieson_brooks:2004,hatfield_etal:2012}.
In order to address these questions, two complicating factors must
be confronted when fitting population models to data. 
First, deterministic models of population dynamics are virtually always inadequate due to process
variation, the inherent stochasticity in demographic parameters and environmental
conditions \citep{bjornstad_grenfell:2001,saether_engen:2002}.
Second, abundance---the natural state variable in studies
of population dynamics---can rarely be observed perfectly in field
studies because of observation error, such as imperfect
detection \citep{link_nichols:1994,kery_etal:2009}.

State-space models are a widely used approach for
studying population dynamics while accounting for process variation
and observation error \citep{devalpine_hastings:2002,
  buckland_etal:2004, dennis_etal:2006}. Classical state-space
models are time-series models in which the true state of the
system (e.g. population size during each year) is observed
imperfectly. One reason for the widespread adoption of
state-space models in ecology is that failure to account for process
variation and observation error can bias estimators of abundance and
population growth parameters. For instance, the strength of
density dependence will be over-estimated if observation error is
ignored \citep{link_nichols:1994,shenk_etal:1998}.

A simple state-space models can be described as follows.
Let $N_t$ be the abundance of a species during year $t$, for
$t=1,\hdots,T$, and let $X_t$ be
the observed data, which differs from $N_t$ due to observation error,
a random effect denoted $\zeta_t$. Temporal variation in $N_t$ is
modeled using a population growth model, $\mu(N_{t-1})$
and random process variation denoted $\grave{o}_t$.
The growth model may be density-dependent, as in the case of the 
Ricker-logistic model, or it might be density-independent, such as when growth
is exponential.
%, in which $\mu(N_{t-1}) = N_{t-1}e^{r}$ where $r$ is the intrinsic rate of increase.
The full model can now be written as:
\begin{subequations}
  \label{eq:ss1}
  \begin{align}
    N_1 &= X_1 \label{eq:ss1a} \\
N_t &= \mu(N_{t-1}) + \grave{o}_{t-1} \quad \text{for} \;
t=2,\hdots,T \label{eq:ss1b} \\
X_t &= N_t + \zeta_t \qquad \qquad \;\, \text{for} \;
t=1,\hdots,T \label{eq:ss1c}
  \end{align}
\end{subequations}
where $\grave{o}_t$ is the random effect allowing for process
variation unaccounted for by the deterministic model. In classical
state-space models, the two sets of random effects
are assumed to be i.i.d. Gaussian deviates:
$\grave{o}_t \sim \mathrm{N}(0, \sigma)$ and
$\zeta_t \sim \mathrm{N}(0, \tau)$. It is also standard practice to
ignore process variation associated with $N_1$, as
indicated by Eq~\ref{eq:ss1a}. Alternatively, the population is
assumed to be at equilibrium such that the $N_1$ can be regarded as an
outcome of the equilibrium distribution.

Even though state-space models such as that shown in Eq.~\ref{eq:ss1}
are among the most widespread approaches for modeling population dynamics
using time-series data, several problems are evident. Namely, (1)
time-series data and the underlying abundance parameters of interest
are typically integer-valued, as in the case of count data, raising
concerns about the use of Gaussian distribution for the random
effects, (2) the model for observation error has little biological basis, (3)
spatial variation in abundance is not allowed, and (4) some of the
parameters of the model are not estimable. We briefly discuss each of
these points before describing a general approach to resolving these
shortcomings.

Use of the Gaussian distribution for modeling random process variation
and observation error is motivated by convenience rather than
biology. Specifically, the Gaussian assumptions
allow for parameter estimation using the Kalman filter
\citep{dennis_etal:2006}, which is much more computationally efficient
than estimation methods when random effects are not Gaussian distributed
\citep{devalpine_hastings:2002}. The problem with this is that it
allows for negative and non-integer values of the state-variable, which is
inconsistent with the observed data, and may result in implausible predictions.
Transforming the count data and the underlying abundance variables do
little to resolve these issues \citep{ohara_kotze:2010}.

Another problem with standard state-space models is that the
observation error has no clear interpretation. The use of a mean zero
Gaussian distribution implies that $X_t$ will be
higher than $N_t$ as often as it is lower than $N_t$. 
It is hard to identify a mechanism that would cause such symmetric errors.
A more likely form of observation error, and one that has been recognized for well
over a century, results from failing to detect individuals that are
present. Imperfect detection may be attributable to
characteristics of the species under study, such as its elusiveness,
or to the failings of the ecologist collecting the data in the field.
Although a vast number of methods have been devised for accounting for
this form of observation error, rarely have these methods been
integrated into state-space models \citep[but
see][]{buckland_etal:2004}.

A more serious problem than the ones associated with the Gaussian
assumptions is that the parameters of the simple state-space models
such as Eq~\ref{eq:ss1} are not identifiable in many
circumstances \citep{polansky_etal:2009}. Recall that the
Eq.~\ref{eq:ss1} did not specify distributions for $N_1$; % or $X_1$;
however, $N_1$ is a random variable and hence there is uncertainty
that should be accounted for.
%are needed to fully specify the model.
To be more specific, a fully-specified state-space model requires at least
three probability distributions, which we represent
using bracket notation (i.e., $[Y|\Omega]$ is interpreted
as the probability distribution of the random variable $Y$ given the
parameter $\Omega$). The three probability distributions required for a
state-space model are:
%%%% JEFF, does this extra sentence help? This notation comes from
%%%% several of the original state-space papers and is widespread in
%%%% books like Andy's and Bill's.
%\hl{[I'm not familiar with bracket notation,
% so these equations are confusing to me. How widely used is this
% notation in ecology? Will others likely understand this? Also, we might want
%to save space by trimming a bunch of this discussion back down.]}
\begin{subequations}
  \label{eq:ss2}
  \begin{align}
    [N_1&|\bm{\theta}] \label{eq:ss2a} \\ \label{eq:ss2b}
[N_t&|N_{t-1},\bm{\Theta}] \quad \text{for} \; t=2,\hdots,T \\
\label{eq:ss2c}
    [X_t&|N_t,\bm{p}]  \qquad \; \text{for} \; t=1,\hdots,T
  \end{align}
\end{subequations}
where $\bm{\theta}$ are the process variation parameters for the
initial state, $\bm{\Theta}$ are the process variation parameters
influencing how abundance changes over time, and $\bm{p}$ are the
observation error parameters. As mentioned previously, Eq.~\ref{eq:ss2b}
and Eq.~\ref{eq:ss2c} are assumed to be Gaussian in classical state-space
models, but what should be the distribution for Eq.~\ref{eq:ss2a}?
%The three lines of general model
%described by Eq.~\ref{eq:ss2} correspond to the three equations shown
%for the specific example in Eq~\ref{eq:ss1}. However, in this case,
%we allow $N_1$ to be a random variable. But what should its
%istribution, $[N_1|\bm{\theta}]$ be?
And how could $\bm{\theta}$ be estimated given that there is only a single observation
available? Often, these problems are simply assumed away. For
example, sometimes researchers assume that there is no process variation in
the first year, although this seems unjustified. Alternatively, the
population may be assumed to be at equilibrium such that the expected
value of $N_t$ is constant through time. Although this 
makes the parameters $\bm{\theta}$ identifiable, assuming equilibrium
defeats the objective of many studies of population dynamics, namely
determining why a population varies over time.

A fourth problem with these models is that they do not admit
spatial variation. This reduces the scope of the inferences that can
be drawn from the models and it ignores the importance of space in
population regulation.

Several extensions of state-space models have been proposed to
overcome the limitations described above. \citet{devalpine_hastings:2002} and
\citet{kery_etal:2009} described methods for fitting models with non-Gaussian
distributions for the process and observation errors. Observation models with
more intuitive interpretations, such as those that explicitly model
detection probability, have been proposed by \citet{kery_etal:2009}. \citet{lele_etal:1998} and
\citet{kery_etal:2009} developed models allowing for inference about
spatial and temporal variation in abundance, and their developments
also resolved the problems of non-identifiability for the parameters
of the initial state at time $t=1$. Of these extensions, only
the work by \citet{kery_etal:2009} addressed several of these limitations
simultaneously. However, their model did not include serial
dependence, which is a hallmark of population models. This
limits the utility of their model for making inferences about explicit
population processes.

In this paper, we focus on the model of \citet[henceforth the DM model]{dail_madsen:2011}
that simultaneous resolves each of the aforementioned problems with
traditional state-space models, and is designed for simple count data.
In the following section, we describe the DM model in its original
form and explain how it resolves each of the deficiencies with
standard state-space models. In Section~\ref{sec:ext}, we extend the
model to accommodate classical models of population growth and
to handle several features common to ecological
time-series. Specifically, we describe methods for accommodating
excess zeros and nuisance variables such as random observer effects.
Both frequentist and Bayesian methods of inference are discussed, and
code for fitting models is presented in the appendices.
%The Bayesian method
%is attractive in that it can accommodate prior information about
%detection parameters, which is available for many species, yet not
%collected as part of many existing monitoring programs.
In Section~\ref{sec:app}, we evaluate the performance of the model
extensions using a simulation study and by analyzing data from the
North American Breeding Bird Survey (BBS), one of
the most spatially and temporally extensive sets of count data on
vertebrate populations \citep{robbins_etal:1986}. The overarching aim
of the paper is to provide ecologists with flexible and accessible means of
addressing important questions related to the variation of abundance in
space and time.
%believe these extensions will increase the utility of the model for
%application to existing datasets such the Breeding Bird Survey.
\begin{comment}
  The DM model allows for inference about spatial and temporal
dynamics in abundance while accounting for observation error
using only spatially and temporally replicated count data. Their model
is an extension of the closed-population $N$-mixture model
\citep{royle:2004biom}, which was designed specifically to
address the problem of modeling spatial variation in abundance in the
  presence of observation error, specifically, when detection
  probability is less than one. The DM extension relaxes the
assumption of population closure and includes explicit
parameters describing population change as a first-order Markovian
  process. [more details]

[applications / potential etc... ] The DM model is relevant to a
vast number of problems confronting basic and applied ecologicalresearch. Answering questions about climate change etc...
require
spatially and temporally extensive datasets such as result from
the
  North American Breeding Bird Survey (BBS)
\citep{robbins_etal:1986}. However, most of the datasets of theappropriate scale are collected by volunteers and using
protocols
  that are not amenable to traditional approaches of modeling
  abundance and detection probability based on standard
capture-recapture methods. Instead they have, at best, simple
count
data for which, when it can be interpreted as counts of unique  individuals, the DM model can be applied.. Yadi ya

Although the DM model was designed explicitly for these
purposes,
the original formulation of the model makes strict
distributional
  assumptions and does not acknowledge many of the sources of
variation inherent to existing ecological time-series data.  We  propose variations on these models that are likely to be more
realistic and useful in many cases. We demonstrate the
usefulness
  of these new models with simulated and real BBS data.

In this paper, we demonstrate how the DM model resolves each ofthese four limitations of conventional state-space models, and
we
extend the DM model in several important ways. First, we
demonstrate
  how classical population models can be embedded within the DM
model. Then we present methods for handling excess zeros as well
as
nuisance parameters such as random observer effects on detection  probability. A simulation study
\end{comment}

\section{The Dail-Madsen Model}
\label{sec:dm}

The DM model is an extension of the $N$-mixture model
\citep{royle:2004biom}, which allows for inference about spatial
variation in abundance when individuals cannot be detected with
certainty. To estimate both abundance parameters and parameters
of the detection process, the original $N$-mixture model uses replicate
observations at each site, which are collected during sufficiently
short time intervals such that the population can safely be
assumed to be closed with respect to births, deaths, and movement. The DM
model relaxes this closure assumption and includes explicit parameters
describing population change over time.

\subsection{The Data}

The DM model requires count data collected at $R$ sites, each of
which is surveyed on $T$ primary sampling periods. A site is ideally a
well-defined region of a study area, such as a wetland or a
patch of early-successional habitat, although it could be an arbitrarily
defined region such as a randomly located survey plots. The time
frame of the study is arbitrary, but in general
it will be sufficiently long to allow for temporal variation in
abundance. For instance, a primary sampling period could be a
one month long breeding season, which could be surveyed once per
year for $T$ years. If the a site is surveyed on $J$ occasions during a
primary period, these are called secondary sampling occasions.
In the case where no secondary sampling was conducted, i.e.
$J=1$, let $X_{it}: i=1,\hdots,R; t=1,\hdots,T$ denote the count data
at site $i$ and primary period $t$. If secondary sampling periods were
used, an extra dimension is added so that we have $X_{ijt}:
j=1,\hdots,J$. Typically, we know $X_{it}$ will be lower than the
actual quantity of interest, abundance,
denoted $N_{it}$. In cases where detection probability is perfect,
abundance is observed directly such that $X_{it} \equiv N_{it}$.

\subsection{The Original Formulation of the Model}

As with traditional state-space models, the DM model includes the
three conditionally related processes shown in Eq.~\ref{eq:ss2},
which correspond to (1) initial abundance, i.e. the
abundance at site $i$ during the first primary period,
denoted (2) abundance at time $t$ (for $t>1$) which depends upon
abundance at $t-1$, and (3) the
detection process \citep{dail_madsen:2011}.
The first two processes describe the state process---the
variation in abundance in space and time. The third process
simply describes the the relationship between
abundance and the observed count data.

\subsubsection{Initial abundance}

Recall that conventional state-space models assume that the
distribution for the initial time period,
$[N_1 | \bm{\theta}]$, is either the equilibrium
distribution, or has zero variance.
In contrast, \citet{dail_madsen:2011} proposed modeling $N_{i1}$
as either a Poisson or negative binomial random variable:
\begin{gather}
N_{i1} \sim \mathrm{Pois}(\Lambda) \nonumber \\
\text{or} \nonumber \\
N_{i1} \sim \mathrm{NB}(\Lambda, \alpha)
\label{eq:N1}
\end{gather}
where $\Lambda_i$ is the expected abundance at site $i$ during
year 1.
The Poisson distribution assumes that the mean of $N_{i1}$ is
equal to its variance, whereas the negative binomial distribution allows the
variance to be greater than the mean with the amount of
overdispersion determined by the parameter $\alpha$.
%In the Poisson case, the equivalent expression of
%$[N_1 | \bm{\theta}]$ (Eq.~\ref{eq:ss2a}) is
%$[N_{i1} | \Lambda] =\frac{\Lambda^{N_{i1}}e^{-\Lambda}}{N_{i1}!}$.
%\begin{equation}
%[N_{i1} | \Lambda] =
%\frac{\Lambda^{N_{i1}}e^{-\Lambda}}{N_{i1}!}.
%  \label{eq:N1br}
%\end{equation}
% Maybe we don't need the Poisson distriubtion equation? - Jeff

Regardless of the specified distribution, the model for initial
abundance has two distinguishing features. First, it provides a
mechanism for characterizing spatial variation in abundance. For
instance, one might consider the influence of some environmental
covariate ($x_i$) on abundance using a log-linear
model such as $\log(\Lambda_i) = \beta^{\Lambda}_0 +
\beta^{\Lambda}_1
x_{i}$. The second important point is that the spatial
replicates resolve the
problem of parameter non-identifiability that are common to
standard state-space models because, as demonstrated by
\citet{royle:2004biom},
process variation and observation error can be estimated from
spatially-replicated count data. Hence, the first component of
the model addresses both the issues of spatial inference and
parameter identifiability discussed previously.

\subsubsection{Abundance in subsequent time periods}

The DM model assumes that abundance in time $t$ is a function of
abundance in time $t-1$, i.e. abundance at each site evolves as a
first order Markovian process, although higher order processes are also possible.
\citet{dail_madsen:2011} considered several models to describe the temporal dynamics;
however, in each case they modeled $N_t$ as the sum of two random variables:
$S_{it}$, the number of indivduals surviving from $t-1$ and not
emigrating; and $G_{it}$ the number of new individuals entering
the population. Their most general model was
\begin{equation}
\left.\begin{aligned}
S_{it}|N_{it-1} &\sim \mathrm{Bin}(N_{it-1}, \omega) \\
G_{it}|N_{it-1} &\sim \mathrm{Pois}(\gamma(N_{it-1})) \\
N_{it} &= S_{it}+G_{it}
\end{aligned}\right\} \quad \text{for} \; t=2,\hdots,T
\label{eq:Nt}
\end{equation}
where $\omega$ is the apparent survival probability and $\gamma$
is the recruitment rate (which can depend on $N_{it-1}$).
\citet{dail_madsen:2011} proposed three
models for $\gamma$: the constant model,
$G_{it} \sim \mathrm{Pois}(\gamma)$ where recruitment does not
depend on $N_{it-1}$, and which simulates a ``propagule rain'' of new
individuals; the autoregressive model, $G_{it} \sim
\mathrm{Pois}(\gamma(N_{it-1}))$, which
simulates geometric or density independent growth; and the
``no-trend'' model, $\gamma = (1-\omega)\Lambda$, which keeps
expected abundance constant over time. As before, covariates of
$\omega$ and $\gamma$ can be easily accommodated, for example
using logit- and log-linear models respectively.

To make the connection between conventional state-space models
and the DM model, we need to replace $[N_t|N_{t-1}, \bm{\Theta}]$,
Eq.~\ref{eq:ss2b}, with an expression derived from
Eq.~\ref{eq:Nt}. This requires summing over all possible values
of $S$ and $G$, which is accomplished using the discrete convolution:
\begin{equation}
[N_{it}|N_{it-1}, \omega, \gamma] =
  \sum_{S_{it-1}=0}^{\min(N_{it},N_{it-1})}
\binom{N_{it-1}}{S_{it-1}}\omega^{S_{it-1}}(1-\omega)^{N_{it-1}-S_{it-1}}
\times
\frac{\gamma^{N_{it}-S_{it-1}}e^{-\gamma}}{(N_{it}-S_{it-1})!}  \label{eq:P}
\end{equation}
where here we assumed no dependence of $\gamma$ on $N_{it-1}$.

In the absence of movement, $\omega$ is exactly the probability of surviving
from year $t$ to $t-1$, and $\gamma$ is the per-capita birth rate
(under the geometric growth model). This ability to directly estimate
demographic parameters from simple count data is one of the DM
model's most appearling features. However,
in the more common scenario when immigration and emigration
occur, $\omega$ and $\gamma$ no longer can be interpreted as vital
rates. Rather, $\omega$ is the probability of surviving and not
emigrating, and $\gamma$ is the sum of the birth rate and the immigration rate,
i.e. the recruitment rate. In such cases, we propose
replacing Eq.~\ref{eq:Nt} with standard population models.

\begin{comment}
Irrespective of the model for population dynamics, it is
worthwhile
to note that the discrete distributions used in the DM model
avoid
the issues associated with the Gaussian distribution used to
model
  process variation in standard state-space models.
\end{comment}

\subsubsection{Observation process}

Eqs.~\ref{eq:N1} and \ref{eq:Nt} fully specify the state model,
i.e. the model for abundance dynamics. This state
model has allowed us to replace the initial abundance
distribution and transition models of conventional state-space models, with
alternatives that (1) allow for modeling spatial variation in
abundance, (2) respect the discrete nature of the counts, and
(3) have parameters that are estimable. The one
remaining task is to replace the observation model
$[X_t|N_t, \bm{p}]$ (Eq.~\ref{eq:ss2c}) with one that can be
more easily interpreted. The observation
model assumes that individuals are missed due to imperfect
detection. The simplest model for imperfect detection is
\begin{equation}
  X_{it} \sim \mathrm{Bin}(N_{it}, p)
  \label{eq:p1}
\end{equation}
%or equivalently,
%\begin{equation}
%[X_{it} | N_{it}, \bm{p}] = \binom{N_{it}}{X_{it}}
%\bm{p}^{X_{it}}(1-\bm{p})^{N_{it}-X_{it}}.
%  \label{eq:1}
%\end{equation}
Under this model, $\bm{p}$ is interpreted as the detection
probability parameter.

%%%% Save for discussion?
%Surprisingly detection probability, $p$, can be estimated without
%secondary sampling occasions. However, this is only possible
%if the parametric assumptions of the population dynamics are met
%exactly. In practice, it is preferable to obtain direct information
%about $p$ using secondary sampling occasions or some other auxiliary
%data, \hl{as is demonstrated in section....}

The entire model can now be written:
\begin{gather}
  N_{i1} \sim \text{Pois}(\Lambda) \nonumber \\
S_{it}|N_{it-1} \sim \mathrm{Bin}(N_{it-1}, \omega) \nonumber \\  
G_{it}|N_{it-1}\sim \mathrm{Pois}(\gamma(N_{it-1})) \\
  N_{it} = S_{it}+G_{it} \nonumber \\
  X_{it} \sim \text{Bin}(N_{it}, p) \nonumber
  \label{eq:hm}
\end{gather}

\section{Statistical Inference}

\subsection{Maximum likelihood}

Maximum likelihood estimation of parameters in models with
discrete random effects requires removing them by
summation. For Markovian models, this can be done recursively,
making the likelihood tractable. A general expression of the likelihood
can be written:
\begin{multline}
  \mathcal{L}(\Lambda, \bm{\Theta}, p | \{X_{it}\}) = \\
  \prod_{i=1}^R \left\{ \sum_{N_{i1}=0}^{\infty}
  [X_{i1}|N_{i1}, p][N_{i1}|\Lambda] \left\{
%  \prod_{t=T}^2 \left\{
  \sum_{N_{i2}=0}^{\infty} \dotsm \sum_{N_{iT}=0}^\infty
  [X_{it}|N_{it}, p][N_{it}|N_{it-1}, \bm{\Theta}] %\right\}
  \right\} \right\}
  \label{eq:like}
\end{multline}
where $\infty$ is replaced by a finite upper bound $N_{max}$, which
%%% Can we rename the upper bound?  We're also using K
%%% for equilibrium abundance.
should be high enough such that the MLEs are not affected by increasing
it further. This likelihood can be maximized numerically using the
\textbf{R} package \texttt{unmarked} \citep{fiske_chandler:2011}. Examples
are given in the Appendix.

\subsection{Bayesian inference}

Bayesian inference has several appealing features. First, it
allows direct probability statements to be made about a hypothesis
given data. [EXPLAIN POSTERIORS AND PRIORS]
Second, Bayesian methods offer straight-forward approaches for
combine data from multiple sources or use existing estimates of parameters as prior
distributions. In practice, simulation methods such as Markov
chain Monte Carlo (MCMC) are used to simulate the posterior
distributions. Although MCMC can be slow, our experience is that it
may actually converge faster than the amount of time required to
maximize the likelihood shown in Eq.\ref{eq:like}, especially
when the upper bound of the summation must be high, e.g. $N_{max}>200$.
Furthermore, MCMC may be the only possible option for estimating parameters
using some of the extensions described below, such as allowing for
additional random effects.

Writeing custom MCMC algorthims is often tedious and foreign to
ecologists, but software packages such as \textbf{WinBUGS} and
\textbf{JAGS} overcome the technical difficulties by allowing
users to specify the model using a simple symbolic descriptions.

%For example here is \textbf{BUGS} code for the ``geometric'' model described
%in Eq.~\ref{eq:hm}:
%\begin{verbatim}
%model {
%lambda ~ dunif(0, 5)
%omega ~ dunif(0,1)
%gamma ~ dunif(0, 10)
%p ~ dunif(0,1)
%for(i in 1:nSites) {
%  N[i,1] ~ dpois(lambda)
%  y[i,1] ~ dbin(p, N[i,1])
%  for(t in 2:nYears) {
%    S[i,t-1] ~ dbin(omega, N[i,t-1])
%    G[i,t-1] ~ dpois(gamma*N[i,t-1])
%    N[i,t] <- S[i,t-1] + G[i,t-1]
%    n[i,t] ~ dbin(p, N[i,t])
%    }
%  }
%}
%\end{verbatim}

\section{Model Extensions}
\label{sec:ext}

\subsection{Population growth models}

% Preliminary DM model runs for several species tended to lead to
% estimates of survival that were unrealistically high and recruitment
% that were unrealistically low, or the reverse (compared to independent
% demographic analyses). The DM models are able to partition changes in
% abundance to survival and recruitment in part by making strong
% distributional assumptions. % (Dail and Madsen 2011).
% When those assumptions are heavily violated the models may proportion
% population growth incorrectly into survival and recruitment, even if
% they estimates population growth accurately.

Although partitioning population growth into survival and recruitment
is ideal in terms of providing a mechanistic description of population
dynamics, it is not always possible using simple count data,
especially when the sites are not closed with respect to immigration
and emigration. %Furthermore, a
%simpler model would have several merits: faster running time, fewer
%total model combinations, and possibly more realistic estimates.
If the mechanistic model must be sacrificed, we suggest replacing it
with classical population growth models that have a long history in
ecology. The simplest of which is the exponential growth
model: $f(N_{it-1}) = N_{it-1}e^r$ where $r$ is the intrinsic
rate of increase. Although demographic and environmental stochasticity
in $r$ could be modeled as functions of covariates or random effects, a
simpler way of allow for stochasticity is to regard $N_{it}$ as a Poisson random variable, simplifying
Eq.~\ref{eq:Nt} to:
\begin{equation}
  N_{it} \sim \text{Pois}(\exp(r)N_{it-1}).
\label{eq:exp}
\end{equation}
This,
like the autoregressive version of the model, is a variant on a simple
density-independent exponential model of population growth.
Density-dependent versions of the model are also possible.  For
example:
\begin{equation}
  N_{it} \sim \text{Pois}(N_{it-1}\exp(r(1-N_{it-1}/K)))
\label{eq:rick}
\end{equation}
where $K$ is the stable equilibrium of the population and $r$ is
the
instantaneous population growth rate at low population
densities, and
both parameters are constrained to be positive. This is a
stochastic version
of the Ricker-logistic population growth model
\citep{ricker:1954}. We also
implemented a modified Gompertz-logistic density-dependent
model \citep{hart_gotelli:2011}:
\begin{equation}
N_{it} \sim
\text{Pois}(N_{it-1}\exp(r(1-\log(N_{it-1}+1)/\log(K+1))))
\label{eq:gomp}
\end{equation}
Here the interpretations of $r$ and $K$ are similar to in the
Ricker-logistic model.

Because a single Poisson distribution controls the
distribution of $N_{it}$ in each of these models, the discrete
convolution used by \citet{dail_madsen:2011} is not required,
speeding up processing time.

\subsection{Immigration models}

The autoregressive, exponential, Ricker-logistic, and Gompertz-logistic
versions of the DM models all share a common feature (or bug):
once the population at a site reaches 0, it must remain at 0.
This is because all contributions to population growth are
local in these models.  We generalized each these models that
include both internal and external (immigration) contributions
to population growth.  The exponential plus immigration
model is:
\begin{equation}
  N_{it} \sim \text{Poisson}(\exp(r)N_{it-1} + \iota)
  \label{eq:expimm2}
\end{equation}
where $\iota$ represents the immigration rate, and is
constrained to be
positive (this is equivalent to separate Poisson processes for
growth and immigration).  This model is close to
the constant DM model (Eq.~\ref{eq:Nt}), with $\exp(r)$ instead
of $\omega$ and $\iota$ instead of $\gamma$, except that the first
process is Poisson distributed instead of binomial. The autoregressive, Ricker-logistic,
and Gompertz-logistic models can be extended to allow for immigration in the
same way.
\begin{comment}
Similarly, the Ricker-logistic + immigration model can be represented as:
the
Gompertz-logistic + immigration model as: (13) and the autoregressive +  immigration model as: (14)
\end{comment}

% Save for discussion
\begin{comment}
These immigration models are based on the assumption that number
of
immigrant is not dependent on local density; however, extendingthese models in this way is conceptually straight-forward, a
Other
  versions of these models with different assumptions are
possible. The Ricker-logistic + immigration and Gompertz-logistic + immigration
models
can be justified by the further assumption that immigrants
arrive as
  adults immediately before the point-count survey (so that
density-dependent processes do not have a chance to act on thenumbers of immigrants; Otto and Day 2007 p. 161).  Deterministicversions of these models do reach stable equilibriums, but theequilibriums are not at K and cannot be solved for analytically  (Otto and Day 2007).
\end{comment}

We have implemented all preceding models in
a maximum likelihood framework by extending the
\texttt{unmarked} package
\citep{fiske_chandler:2011} in \textbf{R} \citep{R-2012}.

%\subsection{Range model of initial abundance}
\subsection{Excess zeros}

\citet{dail_madsen:2011} suggested two distributions for
modeling
initial abundance: Poisson and negative binomial. We have
extended
their model to include another distribution, the zero-inflated
Poisson. This could be useful when, for example, one is modeling
the
abundance of several species with the same set of point
count surveys for all species, but this set includes routes
sites
outside the range of some of the species. The distribution
of
initial abundances can be represented as:
\begin{equation}
N_{i1} \sim \left\{
\begin{aligned}
0 &\; \text{with probability} \; \psi \\
\Lambda &\; \text{with probability} \; (1-\psi)
\end{aligned} \right.
\label{eq:ZIP}
\end{equation}
where $\psi$ represents the proportion of extra zeros.

These models allow three sources of zero counts by observers: a
species was at a site but not detected; the site was within the
species' range but there were no animals at that site in that
year; and the site was outside the species' range. Furthermore,
detection, abundance, and zero-inflation can be modeled separately as
functions of different (or the same) covariates. For example, detection of a
species might depend on wind speed,
abundance on forest type and weather, and zero-inflation upon elevation and
climate.  
%This approach combines elements of occupancy modeling
%\citep{mackenzie_etal:2006} and abundance modeling.
%, and could prove useful
%for distinguishing species' fundamental and realized niches
%(Hutchinson 1957).

%The zero inflation factor can also affect
%recruitment or population dynamics (see below).
%We also considered a zero-inflated negative binomial model of initial
%abundance, but this model did not perform well in preliminary tests.

%\subsection{Dynamic range models}

The zero-inflated Poisson distribution can be applied to not only
initial abundance but also to recruitment and population growth
terms. For example, the recruitment term of the constant DM model
(Eq.~\ref{eq:Nt}) can be modified as follows:
\begin{equation}
G_{it} \sim \left\{
\begin{aligned}
0 &\; \text{with probability} \; \psi \\
\mathrm{Poisson}(\gamma) &\; \text{with probability} \; (1-\psi)\end{aligned} \right.
\label{eq:ZIPts}
\end{equation}
\begin{comment}
When both initial abundance and dynamics are modeled as
zero-inflated,
$\psi$ can be modeled as time-varying or not (in either case,
$\psi$ can also be
modeled as varying over space). In the latter case, it makes
sense
for the model to "remember" which sites are inside the range of
the
species between the initial abundance stage and each time step
of the
dynamics stage (see code sample?).  Adding zero-inflation to the
dynamics of the autoregressive, exponential, Ricker-logistic, or Gompertz-logistic
models
(without immigration) should have no effect when initial
abundance is
also zero-inflated unless $\psi$ is time-varying (because these
dynamic
models cannot recover from zero abundance).
\end{comment}
We have implemented
zero-inflated dynamics in the Bayesian framework using program
\textrm{JAGS} \citep[version 3.2.0]{plummer:2003}.
%(version 3.2.0; Plummer 2003)

\subsection{Environmental and demographic stochasticity}
\begin{comment}
Even though the DM model is conceptually simple, even in
its most stripped down form it includes many random effects (the $N$'s)
making it challenging to compute the likelihood or implement MCMC
algorithms. Nonetheless, in some cases additional random effects
may be of interest. For example, in conventional state-space models,
environmental stochasticity is often modeled as $r_t \sim
\text{Norm}(\mu_r, \sigma_r)$.
\end{comment}
The original DM models include demographic stochasticity (variation in population growth 
rate due to the randomness of birth and death processes) through random 
binomial draws for survival and Poisson draws for recruitment.  Our population
growth extensions use a single Poisson draw for abundance: $N_{it} \sim 
\text{Pois}(f(N_{it-1}))$.  This is also a commonly used (and mathematically equivalent) 
method of modeling demographic stochasticity 
\citep{bonsall_hastings:2004,melbourne_hastings:2008}.

One way to model environmental stochasticity (variation in population growth
rate due to stochastic environmental conditions) is by explicitly
modeling dynamic parameters as functions of environmental covariates, such
as weather.  For example, in the Gompertz-logistic model (Eq.~\ref{eq:gomp}), $r$ could be a 
function of temperature and $K$ a function of rainfall:
\begin{gather}
r_{it} = \exp(r_{0} + r_{temp}T_{it}) \nonumber \\
K_{it} = \exp(K_{0} + K_{rain}R_{it})  \\
N_{it} \sim
\text{Pois}(N_{it-1}\exp(r_{it-1}(1-\log(N_{it-1}+1)/\log(K_{it-1}+1)))) \nonumber
\label{eq:weather}
\end{gather}
where $T_{it}$ is site and year specific temperature, $R_{it}$ is site and year specific rainfall,
and the errors for the coefficients $r_{0}$, $r_{temp}$, $K_{0}$, and $K_{rain}$ are 
normally distributed.

Another way to model environmental stochasticity is as lognormal variation in observed local population 
growth rate \citep{bjornstad:2001,bonsall_hastings:2004}.  This could be applied to both density-independent and density-dependent models.
For example, for the Ricker-logistic + immigration model:
\begin{equation}
N_{it} \sim
\text{Pois}(N_{it-1}\exp(\nu_{it} + r(1-N_{it-1}/K)) + \iota)
\label{eq:nuRand}
\end{equation}
where $\nu$ is a variable with a normal distribution, mean value of 0, and standard deviation $\sigma_\nu$.  
Three variations suggest themselves.  The first has independent environmental stochasticity 
between sites, and might be appropriate where sites are widely dispersed.  In the second 
all sites have the same environmental conditions in a given year (regional stochasticity, $\nu_{it} = \nu_{t}$),
which might be appropriate where the geographic or environmental range is small \citep{hanski:1998}.  
In cases where sites can be close together but the range is broad, a multivariate normal distribution for $\nu_{it}$ with correlation coefficients
between sites in a year dependent on distance (geographic or otherwise) might be appropriate.

\subsection{Random effects of observers}

Variation in detection probability among observers is another
source of random variation that may need to be accounted for applying
the DM model. For example, differences in observers' ability to see,
hear, or identify birds has long been recognized as a potential source of error
in avian point count surveys such as the BBS 
\citep{robbins_etal:1986,sauer_etal:1994auk,diefenbach_etal:2003,campbell_francis:2011}.%,alldredge_etal:2007auk
%Estimating a separate detection probability for each observer
%can be difficult and reduces one's ability to estimate the quantities
%of interest.  This problem is compounded by the fact that
%observers differ greatly in the number of surveys they have
%run (so that many observers' separate detection probabilities
%could not be accurately estimated).

Current BBS trend estimators deal with this problem by
treating observer identity as a random %(as opposed to a fixed)
effect \citep{link_sauer:2002,sauer_link:2011}.
%(Link and Sauer 2002, Sauer and Link 2011).
%This allows observer-specific differences in detection probability,
%but assumes that observers are selected at random from a pool
%of potential observers.  Models that contain both random and
%fixed effects are referred to as mixed models.  Often in mixed
%models (as in this case), the random effect is modeled not
%because it of interest in itself, but to avoid bias in the
%estimates of the fixed effects.
%
To include random observer effects in DM models and these
extensions, Eq.~\ref{eq:p1} can be modified to:
\begin{gather}
X_{ijt} \sim \mathrm{Bin}(N_{it}, p_j) \nonumber \\
\mathrm{logit}(p_j) \sim \mathrm{Normal}(\mu_p, \sigma_p)
\label{eq:pobs}
\end{gather}
where $X_{ijt}$ is the number of bird recorded at site $i$ by
observer $j$ in year $t$, $p_j$ is observer-specific detection probability,
$\mu_p$ is the mean detection probability (on the logit scale), and $\sigma_p$ is
the standard deviation of the random observer effects (also on the logit scale). 
We have implemented random observer effects in the Bayesian
framework using program \textbf{JAGS} \citep[version 3.2.0, see Appendix]{plummer:2003}.
% (version 3.2.0; Plummer 2003)
%with the \textbf{R} package \texttt{rjags} \citep{plummer:2011,R-2012} interface.

\subsection{Forecasting Future Population Size}

State-space models allow for inference about future population size,
a critical component of population viability analysis as well
as conservation planning. Let $N^+_{t}$ denote population size at some
future point in time and $[N^+_t]$ denote the probability
distribution for this random variable. If we can specify this
distribution, we can make statements about the
probability that $N^+_t = k$ where $k$ is any possible
population size including zero.

The classical approach to computing $[N^+_t]$ is based upon
Empirical Bayes methods [CITE]. 
%Note that neither Empiriis not actually a
%Bayesian method, it is used to estim.
%We take the parameter estimates and plug them back in.... For
%example to obtain the posterior distribution for $[N_{i1}]$
%\begin{equation}
%[\hat{N}_{i1}] = \frac{[N_{i1}|\hat{\Lambda}][X_{i1}|N_{i1},
%\hat{p}]}{\sum_{N_{i1}=0}^K
%[N_{i1}|\hat{\Lambda}][X_{i1}|N_{i1},\hat{p}]}
%  \label{eq:eb1}
%\end{equation}
%Once we have estimated the posterior distribution for the first
%time period, we can then proceed to estimate the posterior
%distribution for subsequent time periods using recursion.
%\begin{equation}
%  [\hat{N}_{it}] =
%  \frac{[N_{it}|N_{it-1},\hat{\bm{\Theta}}][X_{it}|N_{it},
%\hat{p}]}{\sum_{N_{it}=0}^K [N_{it}|N_{it-1},
%\hat{\Theta}][X_{it}|N_{it},\hat{p}]}
%  \label{eq:eb2}
%\end{equation}
%
%Note that when $X_{it}$ is not observed, including for future
%time periods, the distribution $[X_{it}|N_{it}]$ is taken to be
%uniform such that it has no influence on the posterior. In other words,
%the posterior for years with no observed data is determined entirely
%by the estimated posterior from the previous year and the
%population dynamics parameters.
%
If covariates of the dynamics parameters are present, the problem
becomes much harder because, to fully account for uncertainty,
we need a model for the future values of the covariates as well. When
the covariates are continuous, this requires integrating over their
possible values, which becomes very challenging
computationally. Indeed, simulation methods such as MCMC are the
only viable options when numerous covariates are present, which will
often be the case in many ecological studies.

Relative to the computational challenges associated with the
empirical Bayes methods, the Bayesian approach is much more easy to
implement, especially since MCMC will be generally be relied upon in the
first place. Furthermore, because population size at any future point
in time is regarded as an unknown parameter like any other, computing the
posterior distribution of $N^+_t$ requires no additional steps
in MCMC algorithm, other than formally stating that the count data are
``missing'' in future years. An example of this is provided in the appendix.

\section{Applications}
\label{sec:app}

\subsection{Simulation Study}

We simulated data for 100 sites over 40 years.  All
simulations assumed initial abundance was Poisson distributed
and no covariates affected initial abundance, dynamics, or
detection probability.  Our first series of simulations
assumed dynamics were exponential.  We ran 1000 simulations for
each combination of low, medium, and high $\Lambda \in
\{1,5,10\}$, $r \in \{-0.005, 0, 0.005\}$, and
$p \in \{0.05, 0.25, 0.5\}$. % (Table 1; 27 total combinations).
Our second series of simulations changed dynamics to the Ricker-logistic
model. We used an initial abundance of 10 and a detection probability
of 0.25, and simulated low, medium, and high values of equilibrium
abundance and maximum growth rate, each with 1000 simulations (Table 1;
9 total combinations). Our third series of simulations was based
on the Ricker-logistic + immigration dynamics model; here we fixed all
parameters the same as the Ricker-logistic model (with r = 0.05 and K = 10) and
simulated low, medium, and high values of immigration rate with 1000
simulation each (Table 1).

We estimated the parameters for each simulation using the same
initial abundance (Poisson) and dynamics models as were
simulated, implemented in the unmarked library in \textbf{R}.
When run in a maximum likelihood framework, these models require a
a finite upper bound $N_{max}$,
\citep{royle:2004biom,dail_madsen:2011};
%(Royle 2004, Dail and Madsen 2011);
we used 200.  We report bias of estimates, root
mean squared error, and coverage (percentage of 95\% confidence
intervals for parameters that overlap the true values).

\begin{comment}
Ideas:
I. Test importance of random effects modeling, and other
differences between JAGS and unmarked
II.	Model with low p values, or a range of p values
III. Ability to distinguish correct initial abundance and
dynamics models
IV.	When can you detect DD if it exists (power)?
V.	Building on the last two: what if your model set doesn't
include the correct DD model? When will the wrong DD model be
favored
over the wrong DID model?
\end{comment}

\subsection{Analysis of Breeding Bird Survey Data}

We applied these models to North American Breeding Bird Survey
(BBS) data from 1966-2010 for two species in the bordering US states
Maryland and Virginia. For our focal species, we selected
ovenbirds (\textit{Seiurus aurocapilla}), an abundant and widespread
forest breeding migrant with a stable or increasing trend in the region \citep{porneluzi_etal:2011}
and golden-winged warblers (\textit{Vermivora chrysoptera}), which
only breeds in the western parts of these states and has been declining in the
region \citep{confer_etal:2011}.  %MORE INFORMATION ABOUT THE SPECIES?

%GENERAL ABOUT BBS DATA.
The BBS is an annual roadside survey implemented by trained
observers in the United States and Canada. An observer conducts 50
3-minute point counts, with 400 m radiuses and 0.8 km apart from each other along a
39.4 km route. We only used data marked as acceptable for use in the annual BBS
analysis \citep{sauer_etal:1994auk}.  We summed the total birds of each
species seen on each route and year and used the routes (rather
than the individual stops) as our sites.
Strong winds can interfere with point count observers' ability
to hear birds \citep{simons_etal:2007}; we tested the effects of wind
speed on detection probability. BBS volunteers record wind conditions
at the beginning and end of each route on the Beaufort Scale
\citep[start and end wind]{robbins_etal:1986}.
%(start and end wind; Robbins et al. 1986)(start and end wind;0-9; cite).
When start or end wind was not recorded we imputed those values with
the start or end mean. We took the mean of start and end wind for
each route and put this average wind scale value into four
categories: 0 $\leq$ wind $<$ 1; 1 $\geq$ wind $<$ 2; 2 $\leq$ wind $<$ 3; and wind
$\geq$ 3 (maximum of 3.5).
Following \citet{link_sauer:2002},
%Link and Sauer (2002),
we also included the first time an observer ran a route as a predictor variable for detection
probability.

%HOW WE OBTAINED AND MANIPULATED PRISM DATA
We ran a series of maximum-likelihood based models for each
species, and then a series of Bayesian models. We started by testing
three models of initial abundance (Poisson, negative binomial, and
zero-inflated Poisson) with exponential dynamics
(Eq.~\ref{eq:exp}) and no covariates.  We selected the minimum Akaike's Information
Criterion (AIC) model from that set to test three additional
models for p: wind, first, and wind + first. We selected the minimum
AIC model from that set to test eight additional models of dynamics:
constant, autoregressive, Ricker-logistic, Gompertz-logistic, autoregressive +
immigration, exponential + immigration, Ricker-logistic +
immigration, and Gompertz-logistic + immigration. %We then tested the effect of average
%minimum temperature for June and July on all dynamics parameters from
%the minimum AIC dynamics model.
As above, these models require a a finite upper bound ($N_{max}$) to integrate
over when run in a maximum likelihood
framework; we used 600 for ovenbirds and 350 for golden-winged warblers.

We ran the top ranked models from the maximum likelihood
analyses in a Bayesian framework.  We added random observer 
effects and, where appropriate, environmental stochasticity or zero-inflated dynamics.  We used
non-informative priors. We tested for lack of convergence using
5 Markov chains for each model \citep{gelman_rubin:1992}.
%(Gelman and Rubin 1992).
For each chain we sampled the MCMC for at least 40,000 iterations, after at
least 3,000 tuning samples.

We also used these models to obtain route and 
year specific abundance estimates.  Because of the memory requirements involved,
we obtained these estimates from a single chain of 5,000 iterations, after 35,000
tuning samples.  We present these estimates for ovenbirds only.
%Latent inclusion factors for temperature effects?

\section{Results}

\subsection{Simulation Study}

Generally, estimator bias and error were low with the exponential model, and 
coverage was nominal (93.4\% - 97.0\%; Fig.~\ref{fig:exp_hists}; Appendix ?). 
However, estimates of $r$ were 
imprecise (RMSE between 0.020 and 0.025; relative RMSE between 413\% and 
482\% for non-zero true $r$) and biased low (between -0.008 and -0.005; relative bias between
-151\% and -101\% for non-zero true $r$) when the true value of $\Lambda$ was low (1). 
Simulations of the Ricker-logistic model also generally performed well (Fig.~\ref{fig:rick_hists}; 
Appendix ?), except when the true value of $r$ was low (0.005). In this case the 
estimates of $r$ and $K$ were inaccurate (relative RMSE between 87.4\% and 2091.8\%),
biased high (between 65.4\% and 198.9\%), and had low coverage (between 77.6\% and 94.3\%).
In other cases for the Ricker-logistic model, relative RMSE was between 6.7\%
and 55\%, bias was between -6.4\% and 5.8\%, and coverage was between
91.9\% and 96.4\%.  All three cases simulated for the Ricker-logistic + immigration model 
performed fairly well (Fig.~\ref{fig:ricki_hists}; Appendix ?; relative RMSE between 7.2\% 
and 62.9\%, bias between -3.5\% and 3.3\%, and coverage between
93.2\% and 96.4\%).

\subsection{Analysis of Breeding Bird Survey Data}

The negative binomial distribution was strongly supported for ovenbird
initial abundance over the Poisson and zero-inflated Poisson
(Table 2A). Even compared to the Poisson, there was
no evidence to support a zero-inflation factor for this species.
The best supported model for p was additive effects of wind speed
and first run (Table 2B, model B.1). First run and increasing wind
speeds both decreased p. All dynamics models with immigration were
better supported than models without immigration (Table 2C); 
the best supported of these was the Ricker-logistic + immigration.
The best supported model of those without immigration was the
Gompertz-logistic, and the autoregressive and the constant model had the least
support.
%There was little support for an effect of minimum June and July
% temperature on dynamics parameters of the Ricker + Immigration model
%(Table 2A, models A.17 - A.18).

Estimates for the top ranked model for ovenbird
%(NB[$\Lambda (.) \alpha (.)$]Ricker-logistic+Immigration[$r(.)K(.) \iota(.)$]p(wind+1st)) 
were similar when
run in the Bayesian framework, except that the estimate of $r$
more than halved (from 0.026 $\pm$ 0.006 to 0.011 $\pm$ 0.005 [for
Bayesian model estimates we present mean and SD]). When random observer effects
were added, estimates for $\Lambda$ and $K$ increased dramatically
(from 31.6 $\pm$ 4.9 to 42.5 $\pm$ 7.3 and from 56.2 $\pm$ 18.2 to 101.8 $\pm$ 21.1,
respectively), and the estimate for r was intermediate (0.018 $\pm$ 0.006). The
estimate of the intercept for $p$ (on the logit scale) dropped from -1.5 $\pm$
0.1 to -2.0 $\pm$ 0.1, and the estimate of $\sigma_p$ was 0.35 $\pm$ 0.03.  When
environmental stochasticity was added, the estimate for $K$ increased (to 112 $\pm$ 58.8) and $r$ was
again intermediate (0.020 $\pm$ 0.010).  Adding environmental stochasticity alone had little effect
on estimate of $p$ or $\Lambda$. The highest estimates of $\Lambda$ and $K$ came 
from the model that included environmental stochasticity and random observer effects (43.8 $\pm$
7.9 and 141.5 $\pm$ 69.5, respectively).  
Gelman and Rubin diagnostics and visual examination of the chain trajectory and density plots
provided some evidence for a lack of convergence in estimates of the
intercept for $p$ for all Bayesian models ran.

We estimated ovenbird abundance by route and year (and averaged over routes) 
using the Bayesian model including random observer effects (Fig.~\ref{fig:oven_N}).  
Estimated average abundance oscillated over time with a slowly increasing trend.  Estimated ovenbird
abundance was highest in western and eastern Maryland in 1970 and 1980, but became
more evenly distributed by 2010.  Coefficient of variation (CV) for route and year specific
abundance estimates varied between 0.11 and 2.31 (mean = 0.46).  By comparison, CV
for the estimate of mean initial abundance ($\Lambda$) was 0.15.

There was slightly more support for the ZIP distribution for initial
abundance of golden-winged warbler than for the negative binomial; both were strongly
supported over the Poisson (Table 3A). The best
supported model for p was an effect of first run (Table 3B, model
B.1). The best supported dynamics model was exponential
(Table 3C, model C.1); the estimate of r from this model was -0.058
$\pm$ 0.017 (SE). There was no support for immigration added to the
exponential or autoregressive models (Table 3C, models C.4 and
C.6).
However, the Gompertz-logistic + immigration and Ricker-logistic + immigration
models were supported over the corresponding models without immigration
(Table 3C, models C.4, C.5, C.7, and C.8), despite low estimates
of $\iota$ from these models (0.003 $\pm$ 0.003 and 0.003 $\pm$
0.002, respectively).  
%There was considerable support for an effect of
%average minimum June and July temperature on r in the population
%growth model (Table 1B, model B.17), with a negative slope.
Because of the closeness of rankings for the ZIP and negative binomial
distributions for golden-winged warbler, we also ran the subsequent models with the
negative binomial initial abundance. Rankings of subsequent
models were similar, but with a p covariate (first) the negative
binomial distribution outranked the ZIP.

Estimates for the top ranked model for golden-winged warbler
%(ZIP[Λ(.)α(.)]Exponential[r(.)]p(1st))
were similar when run in the Bayesian framework. When random
observer effects were added, the estimate for $\Lambda$ increased dramatically
(from 22.4 $\pm$ 7.6 to 55.5 $\pm$ 30.9). The estimate of the intercept
for $p$ (on the logit scale) dropped from -3.5 $\pm$ 0.2 to -5.6 $\pm$ 1.1, and the
estimate of $\sigma_p$ was 1.7 $\pm$ 0.7. Adding environmental stochasticity
also increased the estimate of $\Lambda$ (to 42.9 $\pm$ 17.9) and the model with
both random effects had the highest estimate of $\Lambda$ (74.6 $\pm$ 32.9). 
Adding zero-inflation to the dynamics (as well as
the initial abundance) has no effect for exponential models, because populations
that started at 0 would stay at 0 in any case. Therefore, we tested
zero-inflated dynamics on the Exponential + Immigration model. Adding zero-inflation to the dynamics
increased the estimate of $\iota$ from 0 to 0.091 $\pm$ 0.049, but had only minor
effects on the other parameters. Gelman and Rubin diagnostics and visual
examination of the chain trajectory and density plots provided evidence for a lack of convergence in
estimates of $\Lambda$ for all Bayesian models ran, and for most parameters when random observer
effects were included.

Estimated detection probabilities were generally low. The estimated probability
of detecting an ovenbird in the top model (Table 2, model C.1) varied between 0.157
$\pm$ 0.014 and 0.184 $\pm$ 0.015, depending on wind speed and whether the
observer had run that route before. The estimated probability of detecting a
golden-winged warbler went from 0.013 $\pm$ 0.006 for first time observers to 0.028 $\pm$ 0.008 for
repeat observers for the top model (Table 3, model C.1). Accounting for random
observer effects decreased average detection probability for both species.

For both species, constant dynamics models estimated very high
survival probabilities ($\omega$ = 1 $\pm$ 1.3e-05 for ovenbird and
0.935 $\pm$ 0.011 for
golden-winged warbler). The autoregressive and autoregressive + immigration
models estimated very low survival probabilities for ovenbird ($\omega$ =
0.058 $\pm$ 0.046 and 0.026 $\pm$ 0.060, respectively), 
but not for golden-winged warbler ($\omega$ = 0.715 $\pm$ 0.288 for both models).

\section{Discussion}

\begin{comment}
  The model (1) recognizes
the discrete nature of count data and the underlying true
abundance
state, (2) includes a realistic observation model with explicitdetection probability parameters (3) has estimable parameters,including those describing the initial state at time $t=1$, and
(4)
allows for inference about both temporal and spatial variation
in
  abundance.
\end{comment}

Our work has highlighted the limitations of classical state-space
models as applied to ecological time series data, and we have
demonstrated how a class of open population $N$-mixture models
proposed by \citet{dail_madsen:2011} overcomes these
limitations. We extended their model in several important
ways to accomodate common features of ecological
datasets, including sparse counts, zero-inflation, demographic
stochasticity, and random variation in observation error.
%We made four important developments of the open population
%N-mixture model proposed by \citet{dail_madsen:2011} that make the
%model more applicable to long-term data such as is collected
%by the North American Breeding Bird Survey.
%First, we have
%reconciled the objectives of traditional state-space models
%with open population $N$-mixture models by illustrating how
%classical models of population dynamics can be embedded within
%the framework. Second, we have demonstrated methods of
%accounting for zero-inflation in the time-series. Third, we
%have illustrated how additional random effects such as
%observer-specific detection probabilities can be
%accommodated. Fourth, we presented a Bayesian analysis of the
%model, which, makes it possible to incorporate prior information when
%available, and in some cases, facilitates parameter estimation.
We also demonstrated how many of the objectives of conventional
state-space modeling can be accomplished within this expanded
framework. For example, one of the primary aims of our paper was to demonstrate how classical
population growth models can be embedded in the DM model.
Although we view this as an important connection to make between the two
different classes of state-space models, we believe that
%In some ways,
such population growth models are more phenomenological
than mechanistic, and this runs counter to the motivation for
the DM model. However, estimating
demographic parameters from count data is an ambitious goal, and
the required assumption will not be valid in many cases,
especially when immigration and emigration occur. This was supported by the
unrealistic estimates of vital rates reported here and by the
lack of support for the mechanistic models relative to the population
growth models.

One strategy for adhereing to the original formulation of the
model that included survival and recruitment parameters would be to
use either prior information or additional data. For example, in a
Bayesian analysis existing estimates of demographic parameters
could be used as informative priors, which might make it possible to
seperate the effects of movement versus births and deaths.
Alternatively, ecologists could combine count data with data
from, for example,
capture-recapture data. Models that are fitted to count data and
demographic data are often referred to as integrated population models
\citep[IPM;][]{besbeas_etal:2002,
buckland_etal:2004,schaub_etal:2007}
and we believe that a DM model could work well in this context.
In fact, the DM model could be viewed as a specific case of the
IPM model proposed by \citet{buckland_etal:2004}.
However, we caution that when count data are affected by
movement and vital rates, it would be unwise to ignore the movement
processes.

%In contrast, recent work has sought to use state-space
%models for inference about population processes such as mortality and
%recruitment. To do so, methods have been developed to combine multiple
%sources of information, such as counts and mark-recapture data
%\citep{besbeas_etal:2002, buckland_etal:2004,
%  schaub_etal:2007}.

Although these integrated population models may be viewed as the gold standard in
state-space modeling, ecologists are not always so fortunate to have direct
information about vital rates, especially at large spatial scales.
Rather, count data are much more common, and
are produced by many of the largest monitoring programs in the
world, such as the North American Breeding Bird Survey
\citep[BBS;][]{robbins_etal:1986}.
In the absence of direct information about demographic parameters, and
when the original assumptions of the model do not hold, the
model can still be used for important purposes such as predicting future
population size and changes in species distributions.

Even though the DM models and our extentions are conceptually simple, even in
their most stripped down form they include many random effects (the $N$'s, which
incorporate demographic stochasticity), making it challenging to compute the likelihood 
or implement MCMC algorithms. Nonetheless, in some cases additional random effects
may be of interest.  We have shown a few approaches for incorporating some of these
(environmental stochasticity and random observer effects).  Other approaches to modeling
each of these may also be worth trying.  Incorporating these random effects can increase
the difficulty of achieving model convergence, but they are certainly important.
For example, the estimate of carrying capacity for the ovenbird increased by more than 150\%
when environmental stochasticity and random observer effects were included.

Another important benefit of this class of models is that it can be
readily implemented using frequentist or Bayesian methods by
ecologists lacking skills in computer programming and advanced
statistics. We have included an appendix with extensive code to
simulate and fit many of the extensions discussed in this paper.
In spite of the new extensions we have proposed, several aspects of
the model could be improved. First, the precision with which the
parameters of the state process can be estimated ultimately depends
upon how well detection probability is estimated. When there is
only a single survey per primary period, the information about
detection probability comes from deviations from the parametric
assumptions about population dynamics. Thus, without direct information
about detection probability, the estimates will be determined by the
model's parametric assumptions. Furthermore, properly modeling detection
probability can involve account for nuances such as difference
among individuals and probabilities of being available for detection
\citep{nichols_etal:2009}.

Fortunately, it is straight-forward to incorporate direct information about
detection probability and we recommend that this be done whenever
possible. The original method for proposed in the original paper
was to collected replicated counts within the primary periods during
a period in which the populations could be assumed to be closed. A
robust design could be used to combine multiple surveys per
primary period to increase the precision of the estimates. We envision
that multiple other options are available as well, such as removal,
multiple observer, and distance modeling \citep{williams_etal:2002}. These could be
accomplished by extending the open population $N$-mixture model
in exactly the same way as the closed population version has been
extended \citep[e.g.,][]{royle_etal:2004}.

Our emphasis was on increasing the practical utility of this
class of models, and so we avoided several conceptually-interesting
extensions that we belive would be computationally prohibitive in many
cases. Nonetheless, we will discuss one---spatially-explict models of
immigration. 
In our models immigration is currently modeled as independent of
population sizes in other sites. Several alternatives suggest
themselves if we think about movement in a metapopulation
context \citep{hanski:1998}: immigration depending on total or mean
abundance across sites the previous time step; immigration
depending on abundance in nearby sites only; and immigration
depending on abundance at sites as a function of their distance
from the receiving site \citep{hastings:1991,hanski:1998}. These models would
be based on the assumptions that either the study sites
spatially cover the whole range of the metapopulation or the
annual changes in the populations sampled are representative of
a larger metapopulation. Initial tests suggest that it would be
possible to fit some or all of these models, at least in the
Bayesian framework.
% Could use a better reference than Hastings 1991 for the total abundance 
% immigration models, but haven't been able to find one.  And haven't found
% any references for nearby site immigration (maybe that's because it's not a great idea).
\begin{comment}

For data sources such as the BBS, where stops are nested within
routes, spatially hierarchical versions of these models could be
the best way to model data from individual stops over many
routes. These models would allow the analyst to account for the
potential spatial dependence of stops within a route, by having
the route as another random effect.

Other models of variability in growth or recruitment (such as
negative binomial or generalized Poisson) could be informative.
Stochasticity (represented as variation in the numbers of
animals occurring in a site and year around what the
deterministic dynamics predict) could be over or underdispersed,
compared to the Poisson distribution. Alternatively,
stochasticity could possibly be applied directly to population
model parameters (such as r and K), instead of numbers of
animals.
\end{comment}

%[paragraph on spatially-explicit dispersal]
%[Other possible extensions we could mention:
%Spatially hierarchical model (so can use individual stops)
%Other models of variability in growth or recruitment (such as negative binomial)
%Add density dependence to survival and recruitment models]
%[Discuss causes of low p values (and other BBS results) here or in BBS
%results?]

In our analysis of simulated data, we found that our models generally estimated parameters well.
Two exceptions were the estimates of instantaneous growth rate in exponential model when the 
true value of initial abundance was low and the estimates of equilibrium abundance and maximum
growth rate in the Ricker model when the true value of maximum growth rate was low.  
The failure of the first case was likely due to a large number of sites starting at abundance 0 or reaching it early in the simulation.  
%[Richard, maybe come up with a summary statistic for this based on the simulation N values?]
Since a site that hits abundance 0 in this simulation stayed at 0, this provided little information with
which to estimate growth rate.  The failure of the second case is also reasonable: with very
slow movement toward the equilibrium abundance comes little information for estimating
that abundance or the rate of growth.  In other cases, bias was low and coverage nominal.

We have presented a test case for these models using data from
the BBS for two species. We demonstrate that these models can be
used to model and estimate initial abundance, detection
probability, and density-independent and dependent functions for
population dynamics. We tested these models in both the
frequentist and Bayesian frameworks and found strengths and
weaknesses of each approach. The Bayesian analyses permit more
flexibility and allow the incorporation of random observer
effects and zero-inflated dynamics. Our results suggest ignoring
variation between observers can lead to underestimation of
initial abundance and equilibrium population sizes. On the other
hand, our Bayesian analyses did have convergence problems,
particularly for the species (golden-winged warbler) with sparser data. It's
likely that poor model fits are related to the low estimates of
detection probability, which in turn are probably caused by the
large 400 m radius used for BBS point counts. Again, we believe
that direct information on detectability (such as repeated
visits within a season) would greatly improve the fit of these
models in both frameworks.  We also were able to obtain year
and site-specific estimates of abundance (Fig.~\ref{fig:oven_N}),
although with generally lower precision than more global estimates.  These sort of estimates
may be more reliable when abundance and dynamics are modeled 
as functions of habitat covariates, which we did not do here.
\begin{comment}
[Alternatively, could put discussions of our results where
they fit in by subject area, rather than separately. For
instance, could discuss BBS population dynamics model selection
in population models paragraph, BBS Bayesian vs.
frequentist strengths and weakness in the Bayesian paragraph, 
and golden-winged warbler ZIP results in ZIP
paragraph.]
\end{comment}

The modeling framework we described can be used to address many
pressing issues in ecology and conservation biology. For
example,
%it is possible to test hypotheses about temporal and spatial
%population regulation. Furthermore, the Bayesian approach is very
%useful in that it can be used to combine multiple sources of data to
%develop mechanistic models of population dynamics. Thus one can test
it allows for inference %hypotheses
about the effects of climate change on either explicit
demographic parameters or in derived parameters such as
population growth rate. Furthermore, %under the Bayesian approach, population
in addition to assessing past climate change on population
parameters, projecting populations under future climate scenarios is also
possible. Thus, we expect this modeling methodology can be used
to make quantiative assessments of species vulnerability to climate
change and other important human-induced drivers of population
dynamics.

\section{Acknowledgements}

We thank J. Andrew Royle and T. Scott Sillett for helpful
comments and suggestions. We gratefully acknowledge National Park 
Service, Smithsonian Institution, and [more] for funding support.

\bibliography{Dail_Mad_ext}

\newpage

% Tables

\begin{table}[t]
  \centering
\caption{Changed parameter values, by series of simulations.  We
simulated 1000 sets of data each combination of parameter values
for 100 sites over 40 years.  We assumed that initial abundance
($\Lambda$) was Poisson distributed.  For the exponential 
simulations we included all combinations of low, medium, and
high $\Lambda$, growth rate ($r$), and detection probability ($p$).
For the Ricker-logistic model simulations we used $\Lambda$ = 10 and $p$ = 0.25, and
simulated low, medium, and high values of $r$ and equilibrium density ($K$).
For the Ricker-logistic + immigration dynamics model we fixed all parameters the
same as the Ricker-logistic model (with $r$ = 0.05 and $K$ = 10) and
simulated low, medium, and high values of immigration rate ($\iota$).}  
\begin{tabular}{lcccccccc}
    \hline
    & \multicolumn{3}{c}{Exponential Growth} && \multicolumn{2}{c}{Ricker-logistic} &&
    Ricker-logistic + Immigration \\
    \cline{2-4}     \cline{6-7}    \cline{9-9}
% & \multicolumn{3}{c}{\rule{25mm}{.1pt}} & \rule{5cm}{.5pt} & \multicolumn{2}{c}{\rule{20mm}{.1pt}} \\
& $\Lambda$ & $r$ & $p$ && $r$  & $K$ && $\iota$  \\    
\hline
    Low	        &1	&-0.01	&0.05	&&0.005	 &5	&&0.005  \\
    Med	        &5	&0	&0.25	&&0.05	&10	&&0.05   \\
    High		 &10 &0.005	&0.5	&&0.1	&20	&&0.5    \\
    \hline
  \end{tabular}
\end{table}

\vfill
%\clearpage
\newpage

% Perhaps stack the species
\begin{comment}
\begin{sidewaystable}
  \centering
  \small
\caption{Model selection table for ovenbirds and golden-winged
warblers in Maryland and Virginia, 1966-2010. We present model
name
and number, number of parameters (Par.), and difference in
Akaike's
information criterion between each model and the top model of
that
set ($\Delta$AIC).  For each species, the first section comparesmodels for initial abundance, the second for detection
probability,
    and the third for dynamics.}
  \begin{tabular}[h]{lcclcc}
\hline
A. Ovenbirds	&&&		B. Golden-Winged Warblers \\
\hline
Model	&Par.	&$\Delta$AIC	&Model	&Par.	&$\Delta$AIC \\
A.1. NB[$\Lambda$(.)$\alpha$(.)]Exponential[$r$(.)]$p$(.)	&4	&0
&B.1. ZIP[$\Lambda$(.)$\psi$(.)]Growth[$r$(.)]$p$(.)	&4	&0 \\
A.2. P[$\Lambda$(.)]Growth[$r$(.)]$p$(.)	&3	&1262.7
&B.2. NB[$\Lambda$(.)$\alpha$(.)]Growth[$r$(.)]$p$(.)	&4	&0.4 \\A.3. ZIP[$\Lambda$(.)$\psi$(.)]Growth[$r$(.)]$p$(.)	&4
&1264.7	&B.3. P[$\Lambda$(.)]Growth[$r$(.)]$p$(.)	&3	&37.8 \\
A.4. NB[$\Lambda$(.)$\alpha$(.)]Growth[$r$(.)]$p$(wind+1st)	&8
&0	&B.4. ZIP[$\Lambda$(.)$\psi$(.)]Growth[$r$(.)]$p$(1st)	&5
&0 \\
A.5. NB[$\Lambda$(.)$\alpha$(.)]Growth[$r$(.)]$p$(wind)	&7	&0.9
&B.5. ZIP[$\Lambda$(.)$\psi$(.)]Growth[$r$(.)]$p$(.)	&4	&3.5 \\
A.6. NB[$\Lambda$(.)$\alpha$(.)]Growth[$r$(.)]$p$(1st)	&5	&5.0
&B.6. ZIP[$\Lambda$(.)$\psi$(.)]Growth[$r$(.)]$p$(wind+1st)	&8
&5.9 \\
A.7. NB[$\Lambda$(.)$\alpha$(.)]Growth[$r$(.)]$p$(.)	&4	&6.4
&B.7. ZIP[$\Lambda$(.)$\psi$(.)]Growth[$r$(.)]$p$(wind) &7 &9.3
\\
A.8.
NB[$\Lambda$(.)$\alpha$(.)]Ricker-logistic+Immigration[$r$(.)$K$(.)$\iota$(.)]$p$(wind+1st)
&10	&0	&B.8. ZIP[$\Lambda$(.)$\psi$(.)]Growth[$r$(.)]$p$(1st)
&5	&0 \\
A.9.
NB[$\Lambda$(.)$\alpha$(.)]Gompertz-logistic+Immigration[$r$(.)$K$(.)$\iota$(.)]$p$(wind+1st)
&10	&8.4
&B.9.
ZIP[$\Lambda$(.)$\psi$(.)]Autoregressive[$\gamma$(.)$\omega$(.)]$p$(1st)
&6	&1.7 \\
A.10.
NB[$\Lambda$(.)$\alpha$(.)]Growth+Immigration[$r$(.)$\iota$(.)]$p$(wind+1st)
&9	&36.5
&B.10.
ZIP[$\Lambda$(.)$\psi$(.)]Growth+Immigration[$r$(.)$\iota$(.)]$p$(1st)
&6	&2.0 \\
A.11.
NB[$\Lambda$(.)$\alpha$(.)]Autoregressive+Immigration[$\gamma$(.)$\omega$(.)$\iota$(.)]$p$(wind+1st)
&10	&38.6
&B.11.
ZIP[$\Lambda$(.)$\psi$(.)]Gompertz-logistic+Immigration[$r$(.)$K$(.)$\iota$(.)]$p$(1st)
&7	&2.5 \\
A.12.
NB[$\Lambda$(.)$\alpha$(.)]Gompertz-logistic[$r$(.)$K$(.)]$p$(wind+1st)
&9	&192.8
&B.12. ZIP[$\Lambda$(.)$\psi$(.)]Gompertz-logistic[$r$(.)$K$(.)]$p$(1st)
&6
&3.0 \\
A.13.
NB[$\Lambda$(.)$\alpha$(.)]Ricker-logistic[$r$(.)$K$(.)]$p$(wind+1st)
&9	&195.1
&B.13.
ZIP[$\Lambda$(.)$\psi$(.)]Autoregressive+Immigration[$\gamma$(.)$\omega$(.)$\iota$(.)]$p$(1st)
&7	&3.7 \\
A.14. NB[$\Lambda$(.)$\alpha$(.)]Growth[$r$(.)]$p$(wind+1st)	&8
&271.3
&B.14.
ZIP[$\Lambda$(.)$\psi$(.)]Ricker-logistic+Immigration[$r$(.)$K$(.)$\iota$(.)]$p$(1st)
&7	&4.2 \\
A.15.
NB[$\Lambda$(.)$\alpha$(.)]Autoregressive[$\gamma$(.)$\omega$(.)]$p$(wind+1st)
&9	&273.7
&B.15. ZIP[$\Lambda$(.)$\psi$(.)]Ricker-logistic[$r$(.)$K$(.)]$p$(1st)	&7&5.2 \\
A.16.
NB[$\Lambda$(.)$\alpha$(.)]Constant[$\gamma$(.)$\omega$(.)]$p$(wind+1st)
&9	&1856.7
&B.16.
ZIP[$\Lambda$(.)$\psi$(.)]Constant[$\gamma$(.)$\omega$(.)]$p$(1st)
&6	&12.5 \\
\hline
\end{tabular}
\end{sidewaystable}
\end{comment}

% How about separate tables?  Also updated the Gompertz-logistic results.
\begin{table}
  \centering
  \small
\caption{Model selection table for ovenbirds in Maryland and Virginia,
    1966-2010.  We present model name and number, number of 
parameters (Par.), and difference in Akaike's
information criterion between each model and the top model of
that set ($\Delta$AIC).  The first section compares
models for initial abundance, the second for detection
probability, and the third for dynamics.}
  \begin{tabular}[h]{lcc}
\hline
Model	&Par.	&$\Delta$AIC	\\
\hline
A. Initial Abundance && \\
A.1. NB[$\Lambda$(.)$\alpha$(.)]Exponential[$r$(.)]$p$(.)	&4	&0\\
A.2. P[$\Lambda$(.)]Exponential[$r$(.)]$p$(.)	&3	&1262.7\\
A.3. ZIP[$\Lambda$(.)$\psi$(.)]Exponential[$r$(.)]$p$(.)	&4 &1264.7\\
\hline
B. Detection Probability && \\
B.1. NB[$\Lambda$(.)$\alpha$(.)]Exponential[$r$(.)]$p$(wind+1st)	&8 &0	\\
B.2. NB[$\Lambda$(.)$\alpha$(.)]Exponential[$r$(.)]$p$(wind) &7 &0.9\\
B.3. NB[$\Lambda$(.)$\alpha$(.)]Exponential[$r$(.)]$p$(1st)	&5	&5.0
\\B.4. NB[$\Lambda$(.)$\alpha$(.)]Exponential[$r$(.)]$p$(.)	&4	&6.4\\
\hline
C. Dynamics && \\
C.1. NB[$\Lambda$(.)$\alpha$(.)]Ricker-logistic+Immigration[$r$(.)$K$(.)$\iota$(.)]$p$(wind+1st) &10	&0	\\
C.2. NB[$\Lambda$(.)$\alpha$(.)]Gompertz-logistic+Immigration[$r$(.)$K$(.)$\iota$(.)]$p$(wind+1st) &10	&8.4 \\
C.3. NB[$\Lambda$(.)$\alpha$(.)]Exponential+Immigration[$r$(.)$\iota$(.)]$p$(wind+1st) &9	&36.5\\
C.4. NB[$\Lambda$(.)$\alpha$(.)]Autoregressive+Immigration[$\gamma$(.)$\omega$(.)$\iota$(.)]$p$(wind+1st) &10	&38.6\\
C.5. NB[$\Lambda$(.)$\alpha$(.)]Gompertz-logistic[$r$(.)$K$(.)]$p$(wind+1st) &9	&192.8\\
C.6. NB[$\Lambda$(.)$\alpha$(.)]Ricker-logistic[$r$(.)$K$(.)]$p$(wind+1st) &9	&195.1\\
C.7. NB[$\Lambda$(.)$\alpha$(.)]Exponential[$r$(.)]$p$(wind+1st)	&8 &271.3\\
C.8. NB[$\Lambda$(.)$\alpha$(.)]Autoregressive[$\gamma$(.)$\omega$(.)]$p$(wind+1st) &9	&273.7\\
C.9. NB[$\Lambda$(.)$\alpha$(.)]Constant[$\gamma$(.)$\omega$(.)]$p$(wind+1st) &9	&1856.7\\
\hline
\end{tabular}
\end{table}

\vfill
%\clearpage
\newpage

\begin{table}
  \centering
  \small
  \caption{Model selection table for golden-winged
warblers in Maryland and Virginia, 1966-2010. We present model
name and number, number of parameters (Par.), and difference in
Akaike's information criterion between each model and the top model of
that set ($\Delta$AIC).  The first section compares
models for initial abundance, the second for detection
probability, and the third for dynamics.}
  \begin{tabular}[h]{lcc}
\hline
Model	&Par.	&$\Delta$AIC\\
\hline
A. Initial Abundance && \\
A.1. ZIP[$\Lambda$(.)$\psi$(.)]Exponential[$r$(.)]$p$(.)	&4	&0 \\
A.2. NB[$\Lambda$(.)$\alpha$(.)]Exponential[$r$(.)]$p$(.)	&4	&0.4 \\
A.3. P[$\Lambda$(.)]Exponential[$r$(.)]$p$(.)	&3	&37.8 \\
\hline
B. Detection Probability && \\
B.1. ZIP[$\Lambda$(.)$\psi$(.)]Exponential[$r$(.)]$p$(1st)	&5 &0 \\
B.2. ZIP[$\Lambda$(.)$\psi$(.)]Exponential[$r$(.)]$p$(.)	&4	&3.5 \\
B.3. ZIP[$\Lambda$(.)$\psi$(.)]Exponential[$r$(.)]$p$(wind+1st)	&8 &5.9 \\
B.4. ZIP[$\Lambda$(.)$\psi$(.)]Exponential[$r$(.)]$p$(wind) &7 &9.3\\
\hline
C. Dynamics && \\
C.1. ZIP[$\Lambda$(.)$\psi$(.)]Exponential[$r$(.)]$p$(1st) &5	&0 \\
C.2. ZIP[$\Lambda$(.)$\psi$(.)]Autoregressive[$\gamma$(.)$\omega$(.)]$p$(1st) &6	&1.7 \\
C.3. ZIP[$\Lambda$(.)$\psi$(.)]Exponential+Immigration[$r$(.)$\iota$(.)]$p$(1st) &6	&2.0 \\
C.4. ZIP[$\Lambda$(.)$\psi$(.)]Gompertz-logistic+Immigration[$r$(.)$K$(.)$\iota$(.)]$p$(1st) &7	&2.5 \\
C.5. ZIP[$\Lambda$(.)$\psi$(.)]Gompertz-logistic[$r$(.)$K$(.)]$p$(1st)	&6&3.0 \\
C.6. ZIP[$\Lambda$(.)$\psi$(.)]Autoregressive+Immigration[$\gamma$(.)$\omega$(.)$\iota$(.)]$p$(1st) &7	&3.7 \\
C.7. ZIP[$\Lambda$(.)$\psi$(.)]Ricker-logistic+Immigration[$r$(.)$K$(.)$\iota$(.)]$p$(1st) &7	&4.2 \\
C.8. ZIP[$\Lambda$(.)$\psi$(.)]Ricker-logistic[$r$(.)$K$(.)]$p$(1st)	&7 &5.2 \\
C.9. ZIP[$\Lambda$(.)$\psi$(.)]Constant[$\gamma$(.)$\omega$(.)]$p$(1st) &6	&12.5 \\
\hline
\end{tabular}
\end{table}

\newpage

%%% Figures

\begin{figure}
  \centering
  \includegraphics[height=8in]{figs/exp_hists}
\caption{Histograms of 1000 parameter estimates for each of 27
simulation cases with exponential dynamics. The vertical lines are the 
data-generating values.}
\label{fig:exp_hists}
\end{figure}

\begin{sidewaysfigure}
  \centering
  \includegraphics{figs/rick_hists}
\caption{Histograms of 1000 parameter estimates for each of 9
simulation cases with Ricker-logistic dynamics. The vertical lines are the 
data-generating values.}
\label{fig:rick_hists}
\end{sidewaysfigure}

\begin{sidewaysfigure}
  \centering
  \includegraphics[width=8.5in]{figs/ricki_hists}
\caption{Histograms of 1000 parameter estimates for each of 3
simulation cases with Ricker-logistic + immigration dynamics. The vertical lines are the 
data-generating values.}
\label{fig:ricki_hists}
\end{sidewaysfigure}

\begin{figure}
  \centering
  \includegraphics[width=6.6in]{figs/OVEN_N_by_route_year2}
\caption{Maps of BBS route and year specific estimated abundances for ovenbirds, for the years 1970,
1980, 1990, 2000, and 2010.  Lower right panel depicts estimated mean route abundance by year
(mean and 95\% credible interval).}
\label{fig:oven_N}
\end{figure}

\end{document}