### 2013-09-18 supp/code_samples.tex

\documentclass[12pt]{article}
\usepackage[a4paper,hmargin={2.4cm,2.4cm},vmargin={2.4cm,2.4cm}]{geometry}
\usepackage[mathlines,displaymath]{lineno}
\usepackage{setspace}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{bm}
\usepackage{graphicx}
\usepackage[round,colon,authoryear]{natbib}
\usepackage{color, soul}
\usepackage{fancyvrb}
\usepackage{verbatim}
\usepackage[pdftex,colorlinks=true,citecolor=black,linkcolor=black,urlcolor=black,pdftex,pdfstartview=FitH,bookmarks=true]{hyperref}
\usepackage{rotating}

%%% END Article customizations

%%% The "real" document content comes below...

\begin{document}
\title{Appendix 1: Code Samples}
\maketitle
%\date{} % Activate to display a given date or no date (if empty),
         % otherwise the current date is printed 
\section{\texttt{unmarked} code}

%\subsection{Exponential model}
%\subsection{Ricker-logistic model}
%\subsection{Gompertz-logistic model}
%\subsection{Exponential plus immigration model}
%\subsection{Ricker-logistic plus immigration model}
%\subsection{Gompertz-logistic plus immigration model}
%\subsection{Autoregressive plus immigration model}
%\subsection{Zero-inflated Poisson initial abundance model}
\subsection{Simulation study code}

File sim\_exp.R:
\begin{verbatim}
# Simulate data under exponential growth model
sim.exp <- function(lambda, r, p, nSites=100, nYears=45) {
    y <- N <- matrix(NA, nSites, nYears)
    N[,1] <- rpois(nSites, lambda)
    for(t in 2:nYears) {
        mu <- N[,t-1]*exp(r)
        N[,t] <- rpois(nSites, mu)
    }
    y[] <- rbinom(nSites*nYears, N, p)
    return(list(N=N, y=y, lambda=lambda, r=r, p=p,
                seed=.Random.seed, seed.kind=RNGkind()))
}
\end{verbatim}

File sim\_ricker.r:
\begin{verbatim}
# Simulate data under Ricker model
sim.ricker <- function(lambda, r, K, p, nSites=100, nYears=45) {
    y <- N <- matrix(NA, nSites, nYears)
    N[,1] <- rpois(nSites, lambda)
    for(t in 2:nYears) {
        mu <- N[,t-1]*exp(r*(1-N[,t-1]/K))
        N[,t] <- rpois(nSites, mu)
    }
    y[] <- rbinom(nSites*nYears, N, p)
    return(list(N=N, y=y, lambda=lambda, r=r, p=p,
                seed=.Random.seed, seed.kind=RNGkind()))
}
\end{verbatim}

File sim\_ricki.r:
\begin{verbatim}
# Simulate data under Ricker + immigration model
sim.ricki <- function(lambda, r, K, iota, p, nSites=100, nYears=45) {
    y <- N <- matrix(NA, nSites, nYears)
    N[,1] <- rpois(nSites, lambda)
    for(t in 2:nYears) {
        mu <- N[,t-1]*exp(r*(1-N[,t-1]/K)) + iota
        N[,t] <- rpois(nSites, mu)
    }
    y[] <- rbinom(nSites*nYears, N, p)
    return(list(N=N, y=y, lambda=lambda, r=r, K=K, iota=iota, p=p,
                seed=.Random.seed, seed.kind=RNGkind()))
}
\end{verbatim}

File ricker\_r0.1K5.r:
\begin{verbatim}
# Simulate data under Ricker model

# TO MODIFY
# (1) Change simulator... source(SOMETHING ELSE)
# (2) Change stub
# (3) Change simout
# (4) Change simulation parameters
# (5) Change pcountOpen arguments: "dynamics" and "K"
# (6) Results

ls()
stub <- "ricker_r0.1K5"  # (2)
setwd(paste('simulated', stub, sep='/'))
source("../sim_ricker.r")  # (1)
ls()

library(unmarked)

nsim <- 200
simout.mle <- simout.cover <- matrix(NA, nsim, 4)  # (3)
colnames(simout.mle) <- colnames(simout.cover) <- c("Lambda", "r", "K", "p") # (3)

# Simulate
set.seed(345489)
lambda <- 10 # (4)
r <- 0.1     # (4)
K <- 5       # (4)
p <- 0.25    # (4)
for(i in 1:nsim) {
    cat("\ndoing", i, "at", format(Sys.time()), "\n")
    data.name <- paste("data_", stub, "_", i, sep="")
    data.file <- paste("data/", data.name, ".gzip", sep="")
    sim.i <- sim.ricker(lambda=lambda, r=r, K=K, p=p, nSites=100, nYears=40) # (1)
    cat("   max(N)=", max(sim.i$N), "\n", sep="")
    assign(data.name, sim.i)
    save(list=data.name, file=data.file) # save simulated data
    umf.i <- unmarkedFramePCO(y=sim.i$y, numPrimary=ncol(sim.i$y))
    fm.i <- pcountOpen(~1, ~1, ~1, ~1, umf.i, K=200, dynamics="ricker",  # (5)
                       starts=c(log(lambda), log(r), log(K), qlogis(p))) # (5)
    model.name <- paste("fm_", stub, "_", i, sep="")
    model.file <- paste("fm/", model.name, ".gzip", sep="")
    assign(model.name, fm.i)
    save(list=model.name, file=model.file) # save fitted models
    mle.i <- coef(fm.i)
    simout.mle[i,] <- c(exp(mle.i[1]), exp(mle.i[2]), exp(mle.i[3]), plogis(mle.i[4])) # (6)
    cat("   mle=", round(simout.mle[i,], 4), "\n")
    write.csv(simout.mle, paste(stub, "_mle.csv", sep=""),
              row.names=FALSE)
    CI.lam <- confint(fm.i, type="lambda") # (6)
    CI.r <- confint(fm.i, type="gamma")    # (6)
    CI.K <- confint(fm.i, type="omega")    # (6)
    CI.p <- confint(fm.i, type="det")      # (6)
    simout.cover[i,1] <- log(lambda) >= CI.lam[1] & log(lambda) <= CI.lam[2] # (6)
    simout.cover[i,2] <- log(r) >= CI.r[1] & log(r) <= CI.r[2]               # (6)
    simout.cover[i,3] <- log(K) >= CI.K[1] & log(K) <= CI.K[2]               # (6)
    simout.cover[i,4] <- qlogis(p) >= CI.p[1] & qlogis(p) <= CI.p[2]         # (6)
    write.csv(simout.cover, paste(stub, "_cover.csv", sep=""),
              row.names=FALSE)
    rm(list=c(data.name, model.name))
}
\end{verbatim}

File ricki\_r0.05K10iota0.5.r:
\begin{verbatim}
# Simulate data under Ricker + immigration model

# TO MODIFY
# (1) Change simulator... source(SOMETHING ELSE)
# (2) Change stub
# (3) Change simout
# (4) Change simulation parameters
# (5) Change pcountOpen arguments: "dynamics" and "K"
# (6) Results

stub <- "ricki_r0.05K10iota0.5"  # (2)
setwd(paste('simulated', stub, sep='/'))
ls()
source("../sim_ricki.r")  # (1)
ls()

library(unmarked)

nsim <- 1000
simout.mle <- simout.cover <- matrix(NA, nsim, 5)  # (3)
colnames(simout.mle) <- colnames(simout.cover) <-
    c("Lambda", "r", "K", "iota", "p") # (3)

# Simulate
set.seed(345489)
lambda <- 10 # (4)
r <- 0.05    # (4)
K <- 10      # (4)
iota <- 0.5  # (4)
p <- 0.25    # (4)
for(i in 1:nsim) {
    cat("\ndoing", i, "at", format(Sys.time()), "\n")
    data.name <- paste("data_", stub, "_", i, sep="")
    data.file <- paste("data/", data.name, ".gzip", sep="")
    # Create simulated data
    sim.i <- sim.ricki(lambda=lambda, r=r, K=K, iota=iota, p=p, nSites=100, nYears=40) # (1)
    cat("   max(N)=", max(sim.i$N), "\n", sep="")
    assign(data.name, sim.i)
    # Save simulated data
    save(list=data.name, file=data.file)
    # Run pcountOpen model 
    umf.i <- unmarkedFramePCO(y=sim.i$y, numPrimary=ncol(sim.i$y))
    fm.i <- pcountOpen(~1, ~1, ~1, ~1, umf.i, K=200, dynamics="ricker", # (5)
      immigration=TRUE, starts=c(log(lambda), log(r), log(K), qlogis(p), log(iota))) # (5)
    # Save fitted models
    model.name <- paste("fm_", stub, "_", i, sep="")
    model.file <- paste("fm/", model.name, ".gzip", sep="")
    assign(model.name, fm.i)
    save(list=model.name, file=model.file) 
    # Save maximum likelihood estimates
    mle.i <- coef(fm.i)
    simout.mle[i,] <- c(exp(mle.i[1]), exp(mle.i[2]), exp(mle.i[3]), exp(mle.i[5]), plogis(mle.i[4])) # (6)
    cat("   mle=", round(simout.mle[i,], 4), "\n")
    write.csv(simout.mle, paste(stub, "_mle.csv", sep=""),
              row.names=FALSE)
    # Save coverage
    CI.lam <- confint(fm.i, type="lambda") # (6)
    CI.r <- confint(fm.i, type="gamma")    # (6)
    CI.K <- confint(fm.i, type="omega")    # (6)
    CI.p <- confint(fm.i, type="det")      # (6)
    CI.iota <- confint(fm.i, type="iota")  # (6)
    simout.cover[i,1] <- log(lambda) >= CI.lam[1] & log(lambda) <= CI.lam[2] # (6)
    simout.cover[i,2] <- log(r) >= CI.r[1] & log(r) <= CI.r[2]               # (6)
    simout.cover[i,3] <- log(K) >= CI.K[1] & log(K) <= CI.K[2]               # (6)
    simout.cover[i,4] <- log(iota) >= CI.iota[1] & log(iota) <= CI.iota[2]   # (6)
    simout.cover[i,5] <- qlogis(p) >= CI.p[1] & qlogis(p) <= CI.p[2]         # (6)
    write.csv(simout.cover, paste(stub, "_cover.csv", sep=""),
              row.names=FALSE)
    # Clean up
    rm(list=c(data.name, model.name))
}
\end{verbatim}

\subsection{BBS analysis code}
\begin{verbatim}
library(unmarked)

# Read in input files
ydata <- as.matrix(read.csv('oven3a.csv', row.names=1))
observers <- as.matrix(read.csv('observers.csv', row.names=1))
first.run <- as.matrix(read.csv('first.run.csv', row.names=1))
wind.factor <- as.matrix(read.csv('wind.factor.csv', row.names=1))
# Fix formatting
wind.factor[wind.factor==" 0"] <- "0"
wind.factor[wind.factor==" 1"] <- "1"
wind.factor[wind.factor==" 2"] <- "2"
# Combine obsCovs together into list
obs4 <- list(observers = observers, first.run = first.run, wind.factor = 
  wind.factor)

# Basic settings and labels
year1 <- 66
aou <- 'oven'
gam.model <- list(~1)[[1]] 
om.model <- list(~1)[[1]] 
lam.model <- list(~1)[[1]] 
K <- 600
iota.model <- list(~1) [[1]]

# Input data for pcountOpen
bird.frame <- unmarkedFramePCO(ydata, obsCovs=obs4, numPrimary=45)

# Run mixture models for initial abundance  
mods1 <- rep('', 3)
outputs1 = list()
for (i in 1:3) {
  mixture <- c('P','NB','ZIP')[i]
  dynamics <- c("constant", "autoreg", "trend", "ricker", "gompertz")[3]
  p.model <- list(~1, ~first.run, ~wind.factor, ~wind.factor + first.run)[[1]]
  immigration <- FALSE
  mod.name <- paste(gsub("\\*", ".", gsub(".stand", "", gsub("[ 1]", "", 
    paste(paste(aou, year1, sep=''), mixture, dynamics, 'lam', 
    substring(deparse(lam.model), 2), 'gam', substring(deparse(gam.model), 2), 
    'om', substring(deparse(om.model), 2), 'p', substring(deparse(p.model), 2), 
    ifelse(immigration, "iotaT", "iotaF"), substring(deparse(iota.model), 2),
    'K', sep='.')))), K, sep="")
  mods1[i] <- mod.name
  file.name <- paste(mod.name, 'gzip', sep='.')
  print(file.name)
  fm <- pcountOpen(data=bird.frame, mixture=mixture, gammaformula=gam.model, 
    lambdaformula=lam.model, omegaformula=om.model, pformula=p.model, K=K, 
    dynamics=dynamics, immigration=immigration, iotaformula=iota.model)
  outputs1[[i]] <- fm
  print(summary(fm))
  save(fm, file=file.name)
}
names(outputs1) <- mods1

AIC.table <- function(mods, sort=T) {
  tab <- data.frame(model = names(mods), npar = NA, AIC = NA, delta = NA, 
    weight = NA)
  for (i in 1:length(mods)) {
    tab$npar[i] <- length(coef(mods[[i]]))
    tab$AIC[i] <- mods[[i]]@AIC
  }
  tab$delta <- tab$AIC - min(tab$AIC)
  tab$weight <- exp(-tab$delta/2)/sum(exp(-tab$delta/2))
  if (sort)
    tab <- tab[order(tab$AIC), ]
  return(tab)
}

# Compare these models
mix.aic <- AIC.table(outputs1)
mix.aic
mix.num <- as.integer(rownames(mix.aic)[1])
mix.num

# Run models for detection probability  
mods2 <- c(mods1[mix.num], rep('', 3))
outputs2 <- list(outputs1[[mix.num]])
for (i in 2:4) {
  mixture <- c('P','NB','ZIP')[mix.num]
  dynamics <- c("constant", "autoreg", "trend", "ricker", "gompertz")[3]
  p.model <- list(~1, ~first.run, ~wind.factor, ~wind.factor + first.run)[[i]]
  immigration <- FALSE
  mod.name <- paste(gsub("\\*", ".", gsub(".stand", "", gsub("[ 1]", "", 
    paste(paste(aou, year1, sep=''), mixture, dynamics, 'lam', 
    substring(deparse(lam.model), 2), 'gam', substring(deparse(gam.model), 2), 
    'om', substring(deparse(om.model), 2), 'p', substring(deparse(p.model), 2), 
    ifelse(immigration, "iotaT", "iotaF"), substring(deparse(iota.model), 2),
    'K', sep='.')))), K, sep="")
  mods2[i] <- mod.name
  file.name <- paste(mod.name, 'gzip', sep='.')
  print(file.name)
  fm <- pcountOpen(data=bird.frame, mixture=mixture, gammaformula=gam.model, 
    lambdaformula=lam.model, omegaformula=om.model, pformula=p.model, K=K, 
    dynamics=dynamics, immigration=immigration, iotaformula=iota.model)
  outputs2[[i]] <- fm
  print(summary(fm))
  save(fm, file=file.name)
}
names(outputs2) <- mods2

# Compare these models
p.aic <- AIC.table(outputs2)
p.aic
p.num <- as.integer(rownames(p.aic)[1])
p.num

# Run models for population dynamics  
mods3 <- c(mods2[p.num], rep('', 8))
outputs3 <- list(outputs2[[p.num]])
# Constant dynamics
mixture <- c('P','NB','ZIP')[mix.num]
dynamics <- c("constant", "autoreg", "trend", "ricker", "gompertz")[1]
p.model <- list(~1, ~first.run, ~wind.factor, ~wind.factor + first.run)[[p.num]]
immigration <- FALSE
mod.name <- paste(gsub("\\*", ".", gsub(".stand", "", gsub("[ 1]", "", 
  paste(paste(aou, year1, sep=''), mixture, dynamics, 'lam', 
  substring(deparse(lam.model), 2), 'gam', substring(deparse(gam.model), 2), 
  'om', substring(deparse(om.model), 2), 'p', substring(deparse(p.model), 2), 
  ifelse(immigration, "iotaT", "iotaF"), substring(deparse(iota.model), 2),
  'K', sep='.')))), K, sep="")
mods3[2] <- mod.name
file.name <- paste(mod.name, 'gzip', sep='.')
print(file.name)
fm <- pcountOpen(data=bird.frame, mixture=mixture, gammaformula=gam.model, 
  lambdaformula=lam.model, omegaformula=om.model, pformula=p.model, K=K, 
  dynamics=dynamics, immigration=immigration, iotaformula=iota.model)
outputs3[[2]] <- fm
print(summary(fm))
save(fm, file=file.name)

# Autoregressive dynamics
mixture <- c('P','NB','ZIP')[mix.num]
dynamics <- c("constant", "autoreg", "trend", "ricker", "gompertz")[2]
p.model <- list(~1, ~first.run, ~wind.factor, ~wind.factor + first.run)[[p.num]]
immigration <- FALSE
mod.name <- paste(gsub("\\*", ".", gsub(".stand", "", gsub("[ 1]", "", 
  paste(paste(aou, year1, sep=''), mixture, dynamics, 'lam', 
  substring(deparse(lam.model), 2), 'gam', substring(deparse(gam.model), 2), 
  'om', substring(deparse(om.model), 2), 'p', substring(deparse(p.model), 2), 
  ifelse(immigration, "iotaT", "iotaF"), substring(deparse(iota.model), 2),
  'K', sep='.')))), K, sep="")
mods3[3] <- mod.name
file.name <- paste(mod.name, 'gzip', sep='.')
print(file.name)
fm <- pcountOpen(data=bird.frame, mixture=mixture, gammaformula=gam.model, 
  lambdaformula=lam.model, omegaformula=om.model, pformula=p.model, K=K, 
  dynamics=dynamics, immigration=immigration, iotaformula=iota.model)
outputs3[[3]] <- fm
print(summary(fm))
save(fm, file=file.name)

# Ricker dynamics
mixture <- c('P','NB','ZIP')[mix.num]
dynamics <- c("constant", "autoreg", "trend", "ricker", "gompertz")[4]
p.model <- list(~1, ~first.run, ~wind.factor, ~wind.factor + first.run)[[p.num]]
immigration <- FALSE
mod.name <- paste(gsub("\\*", ".", gsub(".stand", "", gsub("[ 1]", "", 
  paste(paste(aou, year1, sep=''), mixture, dynamics, 'lam', 
  substring(deparse(lam.model), 2), 'gam', substring(deparse(gam.model), 2), 
  'om', substring(deparse(om.model), 2), 'p', substring(deparse(p.model), 2), 
  ifelse(immigration, "iotaT", "iotaF"), substring(deparse(iota.model), 2),
  'K', sep='.')))), K, sep="")
mods3[4] <- mod.name
file.name <- paste(mod.name, 'gzip', sep='.')
print(file.name)
# Ricker and Gompertz generally seem to fail with default starting values, 
# so we set them
starts <- coef(outputs2[[p.num]])
starts <- c(starts[1], -2.35, starts[1]+0.3, starts[3:length(starts)])
fm <- pcountOpen(data=bird.frame, mixture=mixture, gammaformula=gam.model, 
  lambdaformula=lam.model, omegaformula=om.model, pformula=p.model, K=K, 
  dynamics=dynamics, immigration=immigration, iotaformula=iota.model, 
  starts=starts)
outputs3[[4]] <- fm
print(summary(fm))
save(fm, file=file.name)

# Gompertz dynamics
mixture <- c('P','NB','ZIP')[mix.num]
dynamics <- c("constant", "autoreg", "trend", "ricker", "gompertz")[5]
p.model <- list(~1, ~first.run, ~wind.factor, ~wind.factor + first.run)[[p.num]]
immigration <- FALSE
mod.name <- paste(gsub("\\*", ".", gsub(".stand", "", gsub("[ 1]", "", 
  paste(paste(aou, year1, sep=''), mixture, dynamics, 'lam', 
  substring(deparse(lam.model), 2), 'gam', substring(deparse(gam.model), 2), 
  'om', substring(deparse(om.model), 2), 'p', substring(deparse(p.model), 2), 
  ifelse(immigration, "iotaT", "iotaF"), substring(deparse(iota.model), 2),
  'K', sep='.')))), K, sep="")
mods3[5] <- mod.name
file.name <- paste(mod.name, 'gzip', sep='.')
print(file.name)
# Ricker and Gompertz generally seem to fail with default starting values, so we set them
starts <- coef(outputs2[[p.num]])
starts <- c(starts[1], -2.35, starts[1]+0.3, starts[3:length(starts)])
fm <- pcountOpen(data=bird.frame, gammaformula=gam.model, 
  lambdaformula=lam.model, omegaformula=om.model, pformula=p.model, K=K, 
  mixture=mixture, dynamics=dynamics, immigration=immigration, 
  iotaformula=iota.model, starts=starts)
outputs3[[5]] <- fm
print(summary(fm))
save(fm, file=file.name)

# Trend + immigration dynamics
mixture <- c('P','NB','ZIP')[mix.num]
dynamics <- c("constant", "autoreg", "trend", "ricker", "gompertz")[3]
p.model <- list(~1, ~first.run, ~wind.factor, ~wind.factor + first.run)[[p.num]]
immigration <- TRUE
mod.name <- paste(gsub("\\*", ".", gsub(".stand", "", gsub("[ 1]", "", 
  paste(paste(aou, year1, sep=''), mixture, dynamics, 'lam', 
  substring(deparse(lam.model), 2), 'gam', substring(deparse(gam.model), 2), 
  'om', substring(deparse(om.model), 2), 'p', substring(deparse(p.model), 2), 
  ifelse(immigration, "iotaT", "iotaF"), substring(deparse(iota.model), 2),
  'K', sep='.')))), K, sep="")
mods3[6] <- mod.name
file.name <- paste(mod.name, 'gzip', sep='.')
print(file.name)
fm <- pcountOpen(data=bird.frame, mixture=mixture, gammaformula=gam.model, 
  lambdaformula=lam.model, omegaformula=om.model, pformula=p.model, K=K, 
  dynamics=dynamics, immigration=immigration, iotaformula=iota.model)
outputs3[[6]] <- fm
print(summary(fm))
save(fm, file=file.name)

# Autoregressive + immigration dynamics
mixture <- c('P','NB','ZIP')[mix.num]
dynamics <- c("constant", "autoreg", "trend", "ricker", "gompertz")[2]
p.model <- list(~1, ~first.run, ~wind.factor, ~wind.factor + first.run)[[p.num]]
immigration <- TRUE
mod.name <- paste(gsub("\\*", ".", gsub(".stand", "", gsub("[ 1]", "", 
  paste(paste(aou, year1, sep=''), mixture, dynamics, 'lam', 
  substring(deparse(lam.model), 2), 'gam', substring(deparse(gam.model), 2), 
  'om', substring(deparse(om.model), 2), 'p', substring(deparse(p.model), 2), 
  ifelse(immigration, "iotaT", "iotaF"), substring(deparse(iota.model), 2),
  'K', sep='.')))), K, sep="")
mods3[7] <- mod.name
file.name <- paste(mod.name, 'gzip', sep='.')
print(file.name)
fm <- pcountOpen(data=bird.frame, mixture=mixture, gammaformula=gam.model, 
  lambdaformula=lam.model, omegaformula=om.model, pformula=p.model, K=K, 
  dynamics=dynamics, immigration=immigration, iotaformula=iota.model)
outputs3[[7]] <- fm
print(summary(fm))
save(fm, file=file.name)

# Ricker + immigration dynamics
mixture <- c('P','NB','ZIP')[mix.num]
dynamics <- c("constant", "autoreg", "trend", "ricker", "gompertz")[4]
p.model <- list(~1, ~first.run, ~wind.factor, ~wind.factor + first.run)[[p.num]]
immigration <- TRUE
mod.name <- paste(gsub("\\*", ".", gsub(".stand", "", gsub("[ 1]", "", 
  paste(paste(aou, year1, sep=''), mixture, dynamics, 'lam', 
  substring(deparse(lam.model), 2), 'gam', substring(deparse(gam.model), 2), 
  'om', substring(deparse(om.model), 2), 'p', substring(deparse(p.model), 2), 
  ifelse(immigration, "iotaT", "iotaF"), substring(deparse(iota.model), 2),
  'K', sep='.')))), K, sep="")
mods3[8] <- mod.name
file.name <- paste(mod.name, 'gzip', sep='.')
print(file.name)
# Ricker and Gompertz generally seem to fail with default starting values, 
# so we set them
starts <- coef(outputs2[[p.num]])
starts <- c(starts[1], -2.35, starts[1]+0.3, starts[3:(length(starts)-1)], 
  -1.5, starts[length(starts)])
fm <- pcountOpen(data=bird.frame, gammaformula=gam.model, 
  lambdaformula=lam.model, omegaformula=om.model, pformula=p.model, K=K, 
  mixture=mixture, dynamics=dynamics, immigration=immigration, 
  iotaformula=iota.model, starts=starts)
outputs3[[8]] <- fm
print(summary(fm))
save(fm, file=file.name)

# Gompertz + immigration dynamics
mixture <- c('P','NB','ZIP')[mix.num]
dynamics <- c("constant", "autoreg", "trend", "ricker", "gompertz")[5]
p.model <- list(~1, ~first.run, ~wind.factor, ~wind.factor + first.run)[[p.num]]
immigration <- TRUE
mod.name <- paste(gsub("\\*", ".", gsub(".stand", "", gsub("[ 1]", "", 
  paste(paste(aou, year1, sep=''), mixture, dynamics, 'lam', 
  substring(deparse(lam.model), 2), 'gam', substring(deparse(gam.model), 2), 
  'om', substring(deparse(om.model), 2), 'p', substring(deparse(p.model), 2), 
  ifelse(immigration, "iotaT", "iotaF"), substring(deparse(iota.model), 2),
  'K', sep='.')))), K, sep="")
mods3[9] <- mod.name
file.name <- paste(mod.name, 'gzip', sep='.')
print(file.name)
# Ricker and Gompertz generally seem to fail with default starting values, 
# so we set them
starts <- coef(outputs2[[p.num]])
starts <- c(starts[1], -2.35, starts[1]+0.3, starts[3:(length(starts)-1)], 
  -1.5, starts[length(starts)])
fm <- pcountOpen(data=bird.frame, gammaformula=gam.model, 
  lambdaformula=lam.model, omegaformula=om.model, pformula=p.model, K=K, 
  mixture=mixture, dynamics=dynamics, immigration=immigration, 
  iotaformula=iota.model, starts=starts)
outputs3[[9]] <- fm
print(summary(fm))
save(fm, file=file.name)

names(outputs3) <- mods3

# Compare these models
dyn.aic <- AIC.table(outputs3)
dyn.aic
dyn.num <- as.integer(rownames(dyn.aic)[1])
dyn.num
\end{verbatim}

\section{\textbf{JAGS} code}

\subsection{Geometric-recruitment (autoregressive) model}
\begin{verbatim}
model {
lambda ~ dunif(0, 5)
omega ~ dunif(0,1)
gamma ~ dunif(0, 10)
p ~ dunif(0,1)
for(i in 1:nSites) {
  N[i,1] ~ dpois(lambda)
  y[i,1] ~ dbin(p, N[i,1])
  for(t in 2:nYears) {
    S[i,t-1] ~ dbin(omega, N[i,t-1])
    G[i,t-1] ~ dpois(gamma * N[i,t-1])
    N[i,t] <- S[i,t-1] + G[i,t-1]
    y[i,t] ~ dbin(p, N[i,t])
    }
  }
}
\end{verbatim}

\subsection{Exponential model}
\begin{verbatim}
model {
lambda ~ dunif(0, 5)
r ~ dnorm(0,0.01)
p ~ dunif(0,1)
for(i in 1:nSites) {
  N[i,1] ~ dpois(lambda)
  y[i,1] ~ dbin(p, N[i,1])
  for(t in 2:nYears) {
    N[i,t] ~ dpois(exp(r) * N[i,t-1])
    y[i,t] ~ dbin(p, N[i,t])
    }
  }
}
\end{verbatim}

\subsection{Ricker-logistic model}
\begin{verbatim}
model {
lambda ~ dunif(0, 50)
r ~ dunif(0,5)
K ~ dunif(0, 50)
p ~ dunif(0,1)
for(i in 1:nSites) {
  N[i,1] ~ dpois(lambda)
  y[i,1] ~ dbin(p, N[i,1])
  for(t in 2:nYears) {
    muN[i,t-1] <- N[i,t-1] * exp(r * (1 - N[i,t-1] / K))
    N[i,t] ~ dpois(muN[i,t-1])
    y[i,t] ~ dbin(p, N[i,t])
    }
  }
}
\end{verbatim}

\subsection{Gompertz-logistic model}
\begin{verbatim}
model {
lambda ~ dunif(0, 50)
r ~ dunif(0,5)
K ~ dunif(0, 50)
p ~ dunif(0,1)
for(i in 1:nSites) {
  N[i,1] ~ dpois(lambda)
  y[i,1] ~ dbin(p, N[i,1])
  for(t in 2:nYears) {
    muN[i,t-1] <- N[i,t-1] * exp(r * (1 - log(N[i,t-1] + 1) / log(K + 1)))
    N[i,t] ~ dpois(muN[i,t-1])
    y[i,t] ~ dbin(p, N[i,t])
    }
  }
}
\end{verbatim}

\subsection{Exponential plus immigration model}
\begin{verbatim}
model {
lambda ~ dunif(0, 5)
r ~ dnorm(0,0.01)
p ~ dunif(0,1)
iota ~ dunif(0, 5)
for(i in 1:nSites) {
  N[i,1] ~ dpois(lambda)
  y[i,1] ~ dbin(p, N[i,1])
  for(t in 2:nYears) {
    muN[i,t-1] <- N[i,t-1] * exp(r) + iota
    N[i,t] ~ dpois(muN[i,t-1])
    y[i,t] ~ dbin(p, N[i,t])
    }
  }
}
\end{verbatim}

\subsection{Zero-inflated Poisson dynamics model}
\begin{verbatim}
model {
lambda ~ dunif(0, 200)
psi ~ dunif(0, 1)
# Trend + immigration
r ~ dnorm(0, 0.01)
iota ~ dunif(0, 20)
p ~ dunif(0,1)

for(i in 1:nSites) {
  range[i] ~ dbern(1-psi)
  N[i,1] ~ dpois(lambda*range[i])
  for(t in 2:nYears) {
      muN[i,t-1] <- (N[i,t-1]*exp(r)+iota)*range[i]
      N[i,t] ~ dpois(muN[i,t-1])
    }
  for(t in 1:nYears) {
    y[i,t] ~ dbin(p, N[i,t])
    }
  }
}
\end{verbatim}

\subsection{Environmental stochasticity models}
\begin{verbatim}
model {
# Explicitly modeling dynamic parameters as functions of environmental covariates
# Negative binomial initial abundance, Gompertz-logistic dynamics
lambda ~ dunif(0, 400)
alpha ~ dunif(0, 400)
P <- alpha/(alpha + lambda)
K0 ~ dnorm(0, 0.01)
Krain ~ dnorm(0, 0.01)
r0 ~ dnorm(0, 0.01)
rTemp ~ dnorm(0, 0.01)
p ~ dunif(0,1)

for(i in 1:nSites) {
  N[i,1] ~ dnegbin(P, alpha)
  for(t in 2:nYears) {
      K[i,t-1] <- exp(K0 + Krain*rain[i,t-1]) 
      r[i,t-1] <- exp(r0 + rTemp*temp[i,t-1]) 
      muN[i,t-1] <- N[i,t-1] * exp(r[i,t-1] * (1 - log(N[i,t-1] + 1) / log(K[i,t-1] + 1)))
      N[i,t] ~ dpois(muN[i,t-1])
  }
  for(t in 1:nYears) {
    y[i,t] ~ dbin(p, N[i,t])
  }
}
}
\end{verbatim}

\begin{verbatim}
model {
# Independent environmental stochasticity between sites
# Negative binomial initial abundance, Ricker-logistic dynamics + immigration
lambda ~ dunif(0, 400)
alpha ~ dunif(0, 400)
P <- alpha/(alpha + lambda)
K ~ dunif(0, 400)
r ~ dunif(0, 5)
iota ~ dunif(0, 10)
sigma.nu ~ dunif(0, 2)
tau.nu <- 1 / (sigma.nu * sigma.nu)
p ~ dunif(0, 1)

for(i in 1:nSites) {
  N[i,1] ~ dnegbin(P, alpha)
  for(t in 2:nYears) {
      nu[i,t-1] ~ dnorm(0, tau.nu) 
      muN[i,t-1] <- N[i, t-1] * exp(nu[i, t-1] + r * (1 - log(N[i, t-1] + 1) / log(K + 1))) + iota
      N[i,t] ~ dpois(muN[i, t-1])
  }
  for(t in 1:nYears) {
    y[i,t] ~ dbin(p, N[i,t])
  }
}
}
\end{verbatim}

\begin{verbatim}
model {
# Regional stochasticity 
# Negative binomial initial abundance, Ricker-logistic dynamics + immigration
lambda ~ dunif(0, 400)
alpha ~ dunif(0, 400)
P <- alpha/(alpha + lambda)
K ~ dunif(0, 400)
r ~ dunif(0, 5)
iota ~ dunif(0, 10)
sigma.nu ~ dunif(0, 2)
tau.nu <- 1 / (sigma.nu * sigma.nu)
p ~ dunif(0, 1)

for(t in 2:nYears) {
  nu[t-1] ~ dnorm(0, tau.nu)
}

for(i in 1:nSites) {
  N[i,1] ~ dnegbin(P, alpha)
  for(t in 2:nYears) {
      muN[i,t-1] <- N[i, t-1] * exp(nu[t-1] + r * (1 - log(N[i, t-1] + 1) / log(K + 1))) + iota
      N[i,t] ~ dpois(muN[i, t-1])
  }
  for(t in 1:nYears) {
    y[i,t] ~ dbin(p, N[i,t])
  }
}
}
\end{verbatim}

% Don't have this working yet!  Should we leave it out or try to get it working?
%\begin{verbatim}
%model {
%# Correlated environmental stochasticity between sites (exponential decay with distance)
%# Poisson initial abundance, exponential dynamics 
%lambda ~ dunif(0, 100)
%#alpha ~ dunif(0, 100)
%#P <- alpha/(alpha + lambda)
%#K ~ dunif(0, 400)
%r ~ dnorm(0, 0.01)
%#iota ~ dunif(0, 10)
%sigma.nu ~ dunif(0, 2)
%l.nu ~ dunif(0, 1000)
%for(i in 1:nSites) {
%  cor.mat[i,i] <- 1
%#  cov.mat[i,i] <- sigma.nu * sigma.nu
%  for(j in 1:(i-1)) {
%    cor.mat[i,j] <- exp(-((dist.mat[i,j]/(2*l.nu))^2))
%    cor.mat[j,i] <- cor.mat[i,j]
%  }
%}
%cov.mat[1:nSites,1:nSites] <- sigma.nu * cor.mat[1:nSites,1:nSites] * sigma.nu
%tau.nu <- inverse(cov.mat)
%p ~ dunif(0, 1)
%
%
%for(t in 2:nYears) {
%  nu[1:nSites,t-1] ~ dmnorm(mu.nu, tau.nu)
%}
%
%for(i in 1:nSites) {
%  N[i,1] ~ dpois(lambda)
%  for(t in 2:nYears) {
%      muN[i,t-1] <- N[i, t-1] * exp(nu[i, t-1] + r) #+ iota
%      N[i,t] ~ dpois(muN[i, t-1])
%  }
%  for(t in 1:nYears) {
%    y[i,t] ~ dbin(p, N[i,t])
%  }
%}
%}
%
%\end{verbatim}

\subsection{Random effects of observers}
\begin{verbatim}
model {
lambda ~ dunif(0, 50)
r ~ dunif(0,5)
K ~ dunif(0, 50)
p0 ~ dnorm(0, 0.01)
sigma.p ~ dunif(0,10)
tau.p <- 1 / (sigma.p * sigma.p)
p1st ~ dnorm(0, 0.01)

for (o in 1:nObs) {
  logitpObs[o] ~ dnorm(p0,tau.p)
}

for(i in 1:nSites) {
  N[i,1] ~ dpois(lambda)
  logit(p[i,1]) <- logitpObs[obsMat[i,1]] 
  y[i,1] ~ dbin(p[i,1], N[i,1])
  for(t in 2:nYears) {
    muN[i,t-1] <- N[i,t-1] * exp(r * (1 - log(N[i,t-1] + 1) / log(K + 1)))
    N[i,t] ~ dpois(muN[i,t-1])
    logit(p[i,t]) <- logitpObs[obsMat[i,t]] 
    y[i,t] ~ dbin(p[i,t], N[i,t])
    }
  }
}
\end{verbatim}

\subsection{Forecast model}
Since the forecast model can be implemented identically to a purely retrospective model
as far as JAGS code goes, we are including R code as well to show the differences in implementation.
\begin{verbatim}
library(rjags)

# Settings
nRun <- 100000 # Number of times to run the MCMC, after burnin
nThin <- 25 # Thinning rate (only store 1 iteration out of every nThin)
nChain <- 3 # Number of separate chains to run MCMC (for testing convergence)
nBurn <- 10000 # Number of times to run the MCMC before getting output
nAdapt <- 1000 # Number of times to run the MCMC for adaptation at beginning
year1 <- 1966 # First year of estimation data
year2 <- 2010 # Last year of estimation data/first year of projection
year3 <- 2100 # Last year of projection
thresholds <- c(1,10,50) # Quasi-extirpation levels to monitor

# Read in input files
ydata <- as.matrix(read.csv('oven3a.csv', row.names=1))
observers <- as.matrix(read.csv('observers.csv', row.names=1))
wind.factor <- as.matrix(read.csv('wind.factor.csv', row.names=1))
# Fix formatting
wind.factor[wind.factor==" 0"] <- "0"
wind.factor[wind.factor==" 1"] <- "1"
wind.factor[wind.factor==" 2"] <- "2"

# More settings
species = 'oven'
aou <- 6740
nSites <- nrow(ydata)
nYears <- year3 - year1 + 1
nPast <- year2 - year1 + 1
nFuture <- year3 - year2

# File names to save
mod.name <- paste(species, 'sample.proj.JAGS', sep='.')
file.name <- paste(mod.name, 'RData', sep='.')
file.name2 <- paste(mod.name, 'gzip', sep='.')
file.name2

# For projection, fill in the future counts with all NA
ydata <- cbind(ydata, matrix(NA, nrow=nSites, ncol=nFuture))

fakeN <- ydata + 10 # Try to set up plausible initial values for N
fillNAs <- function(N, default) { # Need to replace NAs in initial N with values
  tMax <- ncol(N)
  defaults <- which(is.na(N[,1]) & is.na(N[,2]))
  N[defaults,1] <- default # Assign missing values from the 1st year which also have missing values for the 2nd year with default supplied
  subsequent <- which(is.na(N[,1]))
  N[subsequent,1] <- N[subsequent,2] # Assign missing values from the 1st year which don't have missing values for the 2nd year with the 2nd year's value
  for (t in 2:(tMax-1)) {
    # Assign missing values from this year which also have missing values for the next year with the values from the previous year
    previous <- which(is.na(N[,t]) & is.na(N[,t+1]))
    N[previous,t] <- N[previous,t-1] 
    # Assign missing values from this year which don't have missing values for the next year with the mean of the previous and next year
    means <- which(is.na(N[,t]))
    N[means,t] <- ceiling((N[means, t-1] + N[means, t+1])/2)
  }
  # Assign missing values from the last year with the values from the previous year
  previous <- which(is.na(N[,tMax]))
  N[previous, tMax] <- N[previous, tMax - 1]
  return(N)
}
fakeN <- fillNAs(fakeN, 24)
head(fakeN)

# Replace missing observer numbers with a new value, calculate number of observers
obs.nr5 <- cbind(observers, matrix(NA, nrow=nSites, ncol=nFuture))
nObs <- max(as.vector(obs.nr5), na.rm=T) + 1
obs.nr5[is.na(obs.nr5)] = nObs
# Set wind speed category values, counting routes not run as wind speed 0 
wind1 <- cbind(matrix(as.integer(!is.na(wind.factor) & wind.factor=="1"), nSites, nPast),
  matrix(0, nSites, nFuture))
wind2 <- cbind(matrix(as.integer(!is.na(wind.factor) & wind.factor=="2"), nSites, nPast),
  matrix(0, nSites, nFuture))
wind3 <- cbind(matrix(as.integer(!is.na(wind.factor) & wind.factor=="3+"), nSites, nPast),
  matrix(0, nSites, nFuture))

# Ricker + immigration model, NB initial abundance
# Logit-normal random observer effect
# Wind (factor) effects on p
# Projection model included
sink(file="dm.proj.oven1.txt")
cat("model {
lambda ~ dunif(0, 200) # Average initial abundance
alpha ~ dunif(0, 200) # Initial abundance overdispersion
# JAGS doesn't allow lambda based parameterization of NB, so reparameterize
P <- alpha/(alpha + lambda)
# Carrying capacity (K) 
K ~ dunif(0, 200)
r ~ dunif(0, 2) # Instantaneous growth rate (return to equil. abundance)
iota ~ dunif(0, 10) # Immigration rate
# Detection probability parameters
p0 ~ dnorm(0, 0.01) # Intercept
sigma.p ~ dunif(0,10) # SD due to random observer effects
tau.p <- 1 / (sigma.p * sigma.p) # Convert SD to precision
p1 ~ dnorm(0, 0.01) # Effect of wind speed 1 on detection probability
p2 ~ dnorm(0, 0.01) # Effect of wind speed 2 on detection probability
p3 ~ dnorm(0, 0.01) # Effect of wind speed 3 or higher on detection probability

# Random observer effects, by observer
for (o in 1:nObs) {
  logitpObs[o] ~ dnorm(p0,tau.p)
}

for(i in 1:nSites) {
  N[i,1] ~ dnegbin(P, alpha) #Initial abundance
  # Detection probability 
  logit(p[i,1]) <- logitpObs[obsMat[i,1]] + p1 * wind1[i,1] + p2 * wind2[i,1] + 
    p3 * wind3[i,1] 
  # Observed counts, based on true abundance and detection probabilities
  y[i,1] ~ dbin(p[i,1], N[i,1])
  for(t in 2:nYears) {
    # Expected abundance this time step (Ricker + immigration model)
    muN[i,t-1] <- N[i,t-1] * exp(r * (1 - N[i,t - 1] / K)) + iota
    # Due to demographic stochasticity, actual abundance differs from expected
    N[i,t] ~ dpois(muN[i,t-1])
    # Detection probability 
    logit(p[i,t]) <- logitpObs[obsMat[i,t]] + p1 * wind1[i,t] + 
      p2 * wind2[i,t] + p3 * wind3[i,t]
    # Observed counts, based on true abundance and detection probabilities 
    y[i,t] ~ dbin(p[i,t], N[i,t])
  }
}

# If don't have enough memory to just trace N, instead trace derived variables 
# of interest
for(t in 1:nYears) { 
  meanN[t] <- mean(N[,t]) # Average by year
  for(j in 1:nThresholds) { 
  # Proportion of sites (quasi-)extirpated by year (not cumulative over time)
    PQE[j,t] <- sum(N[,t]<thresholds[j])/nSites 
  } 
}

# If you're also interested in spatial patterns, 
# snapshots for three years, but all sites
N1 <- N[,1] 
N2 <- N[,nPast] 
N3 <- N[,nYears]

}", fill=TRUE)
sink()

# Data to feed into JAGS
dat.proj.oven1 <- list(nSites=nSites, nYears=nYears, nPast=nPast, 
  nObs=nObs, y=ydata, thresholds=thresholds, nThresholds=length(thresholds),
  obsMat = obs.nr5, wind1=wind1, wind2=wind2, wind3=wind3)
# Initial values for parameters
init.proj.oven1 <- function(chain) list(lambda=runif(1, 20, 75), alpha=runif(1, 0, 20), 
  K=runif(1, 10, 50), iota=runif(1), r=runif(1, 0, 0.1), 
  p0=rnorm(1), p1=rnorm(1), p2=rnorm(1), p3=rnorm(1), sigma.p=runif(1),
  N=fakeN, .RNG.name="base::Wichmann-Hill", .RNG.seed=year1 + aou + chain)
# Three options on what parameters to trace
pars.proj.oven1 <- c("lambda", "alpha", "K", "r", "iota", "p0", "p1", "p2", 
  "p3", "sigma.p")
pars.proj.oven1.1 <- c("lambda", "alpha", "K", "r", "iota", "p0", "p1", "p2", 
  "p3", "sigma.p", "meanN", "PQE", "N1", "N2", "N3")
pars.proj.oven1.2 <- c("lambda", "alpha", "K", "r", "iota", "p0", "p1", "p2", 
  "p3", "sigma.p", "N")

set.seed(year1 + aou)
jm.proj.oven1 <- jags.model("dm.proj.oven1.txt", dat.proj.oven1, 
  init.proj.oven1, n.chains=nChain, n.adapt=nAdapt)
update(jm.proj.oven1, n.iter=nBurn)
jc.proj.oven1 <- coda.samples(jm.proj.oven1, pars.proj.oven1.1, n.iter=nRun,
  thin=nThin)
summary(jc.proj.oven1)
save.image(file.name)
png(paste(mod.name, 'converge.\%i.png', sep='.'))
plot(jc.proj.oven1)
dev.off()
gelman.diag(jc.proj.oven1)
save(jc.proj.oven1, file=file.name2)
\end{verbatim}

\subsection{BBS analysis code}
\begin{verbatim}
# Read in input files
ydata <- as.matrix(read.csv('oven3a.csv', row.names=1))
obs5 <- as.matrix(read.csv('observers.csv', row.names=1))
first <- as.matrix(read.csv('first.run.csv', row.names=1))
wind.factor <- as.matrix(read.csv('wind.factor.csv', row.names=1))
# Fix formatting
wind.factor[wind.factor==" 0"] <- "0"
wind.factor[wind.factor==" 1"] <- "1"
wind.factor[wind.factor==" 2"] <- "2"

library(rjags)

# Settings
year1 <- 66
species <- 'oven'
nSites <- nrow(ydata)
nYears <- ncol(ydata)

fillN3 <- function(N, p, lambda, alpha) { # Need to replace NAs in initial N with values, adjust others by detection probability
  for (i in 1:nrow(N)) {
    notNA <- which(!is.na(N[i, ]))
    size <- N[i, notNA] 
    size[size==0] <- 0.5
    addTo <- rnbinom(length(notNA), size=size, prob=p)
    N[i, notNA] <- N[i, notNA] + addTo
    for (t in 1:ncol(N))
      if (is.na(N[i,t])) {
        if (t == ncol(N) || is.na(N[i, t+1])) {
          if (t==1) {
            N[i, t] <- rnbinom(1, size=alpha, mu=lambda) #If no previous or subsequent value, use supplied default
          }
          else {
            N[i, t] <- N[i, t-1] #If no subsequent value, use previous
          }
        }
        else if (t==1)
          N[i, t] <- N[i, t+1] #If no previous value, use subsequent
        else
          N[i, t] <- ceiling((N[i, t-1] + N[i, t+1])/2) #Otherwise, use mean
      }
  }
  return(N)
}

N <- fillN3(ydata, 0.18, 33.4, 0.65)
head(N)
summary(N)

# Other input variables
nObs <- max(as.vector(obs5), na.rm=T) + 1
obs5[is.na(obs5)] <- nObs
wind1 <- matrix(as.integer(!is.na(wind.factor) & wind.factor=="1"), nSites, nYears)
summary(wind1)
wind2 <- matrix(as.integer(!is.na(wind.factor) & wind.factor=="2"), nSites, nYears)
wind3 <- matrix(as.integer(!is.na(wind.factor) & wind.factor=="3+"), nSites, nYears)
first[is.na(first)] <- 1

# Repeat top model from frequentist analysis
# Pick filename based on model
mixture <- c('P','NB','ZIP')[2]
dynamics <- c("constant", "autoreg", "trend", "ricker", "gompertz")[4]
p.model <- list(~1, ~first.run, ~wind.factor, ~wind.factor + first.run)[[4]] 
r.model <- list(~1)[[1]] 
K.model <- list(~1)[[1]] 
lam.model <- list(~1)[[1]] 
immigration <- TRUE
iota.model <- list(~1) [[1]]
p.rand <- c("None","Normal","Beta")[1]
mod.name <- paste(paste(species, year1, sep=''), mixture, dynamics, 'lam', 
  substring(deparse(lam.model), 2), 'r', substring(deparse(r.model), 2), 
  'K', substring(deparse(K.model), 2), 
  ifelse(immigration, "iotaT", "iotaF"), substring(deparse(iota.model), 2),
  'p', substring(deparse(p.model), 2), 'rand', p.rand, 'JAGS', sep='.') 
file.name <- paste(mod.name, 'RData', sep='.')
file.name

sink(file="dm.ricki1.txt")
cat("
model {
lambda ~ dunif(0, 200)
alpha ~ dunif(0, 200)
p.nb <- alpha/(alpha + lambda)
r ~ dunif(0, 5)
K ~ dunif(0, 200)
p0 ~ dnorm(0, 0.01)
p1 ~ dnorm(0, 0.01)
p2 ~ dnorm(0, 0.01)
p3 ~ dnorm(0, 0.01)
p1st ~ dnorm(0, 0.01)
iota ~ dunif(0, 15)
for(i in 1:nSites) {
  N[i,1] ~ dnegbin(p.nb, alpha)
  for(t in 2:nYears) {
      muN[i,t-1] <- N[i,t-1]*exp(r*(1-N[i,t-1]/K))+iota
      N[i,t] ~ dpois(muN[i,t-1])
    }
  for(t in 1:nYears) {
    logit(p[i,t]) <- p0 + p1*wind1[i,t] + p2*wind2[i,t] + p3*wind3[i,t] + p1st*first[i,t]
    y[i,t] ~ dbin(p[i,t], N[i,t])
    }
  }
}
", fill=TRUE)
sink()

dat.ricki1 <- list(nSites=nSites, nYears=nYears, y=ydata, wind1=wind1, 
  wind2=wind2, wind3=wind3, first=first)
init.ricki1 <- function() list(lambda=runif(1, 10, 60), alpha=runif(1, 0, 60), 
  r=runif(1), K=runif(1, 30, 120), iota=runif(1),
  p0=rnorm(1, -1.5), p1=rnorm(1), p2=rnorm(1), p3=rnorm(1), p1st=rnorm(1), N=N)
pars.ricki1 <- c("lambda", "alpha", "r", "K", "p0", "p1", "p2", "p3", "p1st", "iota")

jm.ricki1 <- jags.model("dm.ricki1.txt", dat.ricki1, init.ricki1, n.chains=5, n.adapt=5000)
jc.ricki1 <- coda.samples(jm.ricki1, pars.ricki1, n.iter=100000)

summary(jc.ricki1)
windows(record=T)
plot(jc.ricki1)
warnings()
gelman.diag(jc.ricki1)

jc.ricki1a <- window(jc.ricki1, 51001)
summary(jc.ricki1a)
gelman.diag(jc.ricki1a)

save.image(file.name)

# Add random observer effects
rm(jc.ricki1)
rm(jc.ricki1a)
p.rand <- c("None","Normal","Beta")[2]
mod.name <- paste(paste(species, year1, sep=''), mixture, dynamics, 'lam', 
  substring(deparse(lam.model), 2), 'r', substring(deparse(r.model), 2), 
  'K', substring(deparse(K.model), 2), 
  ifelse(immigration, "iotaT", "iotaF"), substring(deparse(iota.model), 2),
  'p', substring(deparse(p.model), 2), 'rand', p.rand, 'JAGS', sep='.') 
file.name = paste(mod.name, 'RData', sep='.')
file.name

sink(file="dm.ricki2.txt")
cat("
model {
lambda ~ dunif(0, 200)
alpha ~ dunif(0, 200)
p.nb <- alpha/(alpha + lambda)
r ~ dunif(0, 5)
K ~ dunif(0, 200)
p0 ~ dnorm(0, 0.01)
p1 ~ dnorm(0, 0.01)
p2 ~ dnorm(0, 0.01)
p3 ~ dnorm(0, 0.01)
p1st ~ dnorm(0, 0.01)
iota ~ dunif(0, 15)
sigma.p ~ dunif(0,15)
tau.p <- 1 / (sigma.p * sigma.p)

for (o in 1:nObs) {
  eta[o] ~ dnorm(0,tau.p)
}

for(i in 1:nSites) {
  N[i,1] ~ dnegbin(p.nb, alpha)
  for(t in 2:nYears) {
      muN[i,t-1] <- N[i,t-1]*exp(r*(1-N[i,t-1]/K))+iota
      N[i,t] ~ dpois(muN[i,t-1])
    }
  for(t in 1:nYears) {
    logit(p[i,t]) <- p0 + p1*wind1[i,t] + p2*wind2[i,t] + p3*wind3[i,t] + p1st*first[i,t] + eta[obsID[i,t]]
    y[i,t] ~ dbin(p[i,t], N[i,t])
    }
  }
}
", fill=TRUE)
sink()

dat.ricki2 <- list(nSites=nSites, nYears=nYears, y=ydata, wind1=wind1, 
  wind2=wind2, wind3=wind3, first=first, nObs=nObs, obsID=obs5)
init.ricki2 <- function() list(lambda=runif(1, 10, 60), alpha=runif(1, 0, 60), 
  r=runif(1, 0, 0.1), K=runif(1, 40, 70),
  p0=rnorm(1, -1.5), p1=rnorm(1), p2=rnorm(1), p3=rnorm(1), p1st=rnorm(1), 
  iota=runif(1), sigma.p=runif(1, 0, 2), N=N)
pars.ricki2 <- c("lambda", "alpha", "r", "K", "iota", "p0", "p1", "p2", "p3", 
  "p1st", "sigma.p")

jm.ricki2 <- jags.model("dm.ricki2.txt", dat.ricki2, init.ricki2, n.chains=5, n.adapt=3000)
jc.ricki2 <- coda.samples(jm.ricki2, pars.ricki2, n.iter=40000)

summary(jc.ricki2)
windows(record=T)
plot(jc.ricki2)
warnings()

save.image(file.name)

# Regional environmental stochasticity
rm(jc.ricki2)
p.rand <- c("None","Normal","Beta")[1]
mod.name <- paste(paste(species, year1, sep=''), mixture, dynamics, 'lam', 
  substring(deparse(lam.model), 2), 'r', substring(deparse(r.model), 2), 
  'K', substring(deparse(K.model), 2), 
  ifelse(immigration, "iotaT", "iotaF"), substring(deparse(iota.model), 2),
  'p', substring(deparse(p.model), 2), 'rand', p.rand, 'JAGS.ES', sep='.') 
file.name = paste(mod.name, 'RData', sep='.')
file.name

sink(file="dm.ricki3.txt")
cat("
model {
lambda ~ dunif(0, 400)
alpha ~ dunif(0, 200)
p.nb <- alpha/(alpha + lambda)
r ~ dunif(0, 3)
sigma.nu ~ dunif(0, 10)
tau.nu <- 1 / (sigma.nu * sigma.nu)
K ~ dunif(0, 400)
p0 ~ dnorm(0, 0.01)
p1 ~ dnorm(0, 0.01)
p2 ~ dnorm(0, 0.01)
p3 ~ dnorm(0, 0.01)
p1st ~ dnorm(0, 0.01)
iota ~ dunif(0, 15)

for(t in 2:nYears) {
  nu[t-1] ~ dnorm(0, tau.nu)
}

for(i in 1:nSites) {
  N[i,1] ~ dnegbin(p.nb, alpha)
  for(t in 2:nYears) {
      muN[i,t-1] <- N[i,t-1]*exp(nu[t-1]+r*(1-N[i,t-1]/K))+iota
      N[i,t] ~ dpois(muN[i,t-1])
    }
  for(t in 1:nYears) {
    logit(p[i,t]) <- p0 + p1*wind1[i,t] + p2*wind2[i,t] + p3*wind3[i,t] + p1st*first[i,t]
    y[i,t] ~ dbin(p[i,t], N[i,t])
    }
  }
}
", fill=TRUE)
sink()

dat.ricki3 <- list(nSites=nSites, nYears=nYears, y=ydata, wind1=wind1, 
  wind2=wind2, wind3=wind3, first=first) 
init.ricki3 <- function() list(lambda=runif(1, 10, 60), alpha=runif(1, 0, 60), 
  r=runif(1, 0, 0.2), sigma.nu=runif(1), K=runif(1, 30, 120), iota=runif(1),
  p0=rnorm(1, -1.5), p1=rnorm(1), p2=rnorm(1), p3=rnorm(1), p1st=rnorm(1), N=N)
pars.ricki3 <- c("lambda", "alpha", "r", "sigma.nu", "K", "iota", 
  "p0", "p1", "p2", "p3", "p1st")

jm.ricki3 <- jags.model("dm.ricki3.txt", dat.ricki3, init.ricki3, n.chains=5, n.adapt=1000)
update(jm.ricki3, 50000)
jc.ricki3 <- coda.samples(jm.ricki3, pars.ricki3, n.iter=130000)

summary(jc.ricki3)
save.image(file.name)
HPDinterval(jc.ricki3)
gelman.diag(jc.ricki3)
png(paste(mod.name, '%i.png', sep='.'))
plot(jc.ricki3)
dev.off()

# Regional environmental stochasticity AND random observer effects
summary(jc.ricki3)
p.rand <- c("None","Normal","Beta")[2]
mod.name <- paste(paste(species, year1, sep=''), mixture, dynamics, 'lam', 
  substring(deparse(lam.model), 2), 'r', substring(deparse(r.model), 2), 
  'K', substring(deparse(K.model), 2), 
  ifelse(immigration, "iotaT", "iotaF"), substring(deparse(iota.model), 2),
  'p', substring(deparse(p.model), 2), 'rand', p.rand, 'JAGS.ES', sep='.') 
file.name <- paste(mod.name, 'RData', sep='.')
file.name

sink(file="dm.ricki4.txt")
cat("
model {
lambda ~ dunif(0, 400)
alpha ~ dunif(0, 200)
p.nb <- alpha/(alpha + lambda)
r ~ dunif(0, 3)
sigma.nu ~ dunif(0, 10)
tau.nu <- 1 / (sigma.nu * sigma.nu)
K ~ dunif(0, 1000)
p0 ~ dnorm(0, 0.01)
p1 ~ dnorm(0, 0.01)
p2 ~ dnorm(0, 0.01)
p3 ~ dnorm(0, 0.01)
p1st ~ dnorm(0, 0.01)
iota ~ dunif(0, 15)
sigma.p ~ dunif(0,15)
tau.p <- 1 / (sigma.p * sigma.p)

for (o in 1:nObs) {
  eta[o] ~ dnorm(0,tau.p)
}

for(t in 2:nYears) {
  nu[t-1] ~ dnorm(0, tau.nu)
}

for(i in 1:nSites) {
  N[i,1] ~ dnegbin(p.nb, alpha)
  for(t in 2:nYears) {
      muN[i,t-1] <- N[i,t-1]*exp(nu[t-1]+r*(1-N[i,t-1]/K))+iota
      N[i,t] ~ dpois(muN[i,t-1])
    }
  for(t in 1:nYears) {
    logit(p[i,t]) <- p0 + p1*wind1[i,t] + p2*wind2[i,t] + p3*wind3[i,t] + p1st*first[i,t] + eta[obsID[i,t]]
    y[i,t] ~ dbin(p[i,t], N[i,t])
    }
  }
}
", fill=TRUE)
sink()

dat.ricki4 <- list(nSites=nSites, nYears=nYears, y=ydata, wind1=wind1, 
  wind2=wind2, wind3=wind3, nObs=nObs, obsID=obs5, first=first)
init.ricki4 <- function() list(lambda=runif(1, 10, 60), alpha=runif(1, 0, 60), 
  r=runif(1, 0, 0.2), sigma.nu=runif(1), K=runif(1, 30, 120), iota=runif(1),
  p0=rnorm(1, -1.5), p1=rnorm(1), p2=rnorm(1), p3=rnorm(1), p1st=rnorm(1), sigma.p=runif(1), N=N)
pars.ricki4 <- c("lambda", "alpha", "r", "sigma.nu", "K", "iota", 
  "p0", "p1", "p2", "p3", "p1st", "sigma.p")

jm.ricki4 <- jags.model("dm.ricki4.txt", dat.ricki4, init.ricki4, n.chains=5, n.adapt=1000)
update(jm.ricki4, 50000)
jc.ricki4 <- coda.samples(jm.ricki4, pars.ricki4, n.iter=130000)

summary(jc.ricki4)
save.image(file.name)
HPDinterval(jc.ricki4)
gelman.diag(jc.ricki4)

jpeg(paste(mod.name, '%i.jpg', sep='.'))
plot(jc.ricki4)
dev.off()

# Track population sizes (includes random observer effects but not environmental stochasticity)
rm(jc.ricki4)
p.rand <- c("None","Normal","Beta")[2]
mod.name <- paste(paste(species, year1, sep=''), mixture, dynamics, 'lam', 
  substring(deparse(lam.model), 2), 'r', substring(deparse(r.model), 2), 
  'K', substring(deparse(K.model), 2), 
  ifelse(immigration, "iotaT", "iotaF"), substring(deparse(iota.model), 2),
  'p', substring(deparse(p.model), 2), 'rand', p.rand, 'JAGS.N', sep='.') 
file.name <- paste(mod.name, 'RData', sep='.')
file.name

sink(file="dm.ricki5.txt")
cat("
var Ndecade[nSites,5]
model {
lambda ~ dunif(0, 200)
alpha ~ dunif(0, 200)
p.nb <- alpha/(alpha + lambda)
r ~ dunif(0, 5)
K ~ dunif(0, 200)
p0 ~ dnorm(0, 0.01)
p1 ~ dnorm(0, 0.01)
p2 ~ dnorm(0, 0.01)
p3 ~ dnorm(0, 0.01)
p1st ~ dnorm(0, 0.01)
iota ~ dunif(0, 15)
sigma.p ~ dunif(0,15)
tau.p <- 1 / (sigma.p * sigma.p)

for (o in 1:nObs) {
  eta[o] ~ dnorm(0,tau.p)
}

for(i in 1:nSites) {
  N[i,1] ~ dnegbin(p.nb, alpha)
  for(t in 2:nYears) {
      muN[i,t-1] <- N[i,t-1]*exp(r*(1-N[i,t-1]/K))+iota
      N[i,t] ~ dpois(muN[i,t-1])
    }
  for(t in 1:nYears) {
    logit(p[i,t]) <- p0 + p1*wind1[i,t] + p2*wind2[i,t] + p3*wind3[i,t] + p1st*first[i,t] + eta[obsID[i,t]]
    y[i,t] ~ dbin(p[i,t], N[i,t])
    }
  }

for(t in 1:nYears) {
  Nmean[t] <- mean(N[,t])
}

for (t2 in 1:length(decades)) {
  for(i in 1:nSites) {
    Ndecade[i,t2] <- N[i,decades[t2]]
  }
}
}
", fill=TRUE)
sink()

dat.ricki5 <- list(nSites=nSites, nYears=nYears, y=ydata, wind1=wind1, 
  wind2=wind2, wind3=wind3, nObs=nObs, obsID=obs5, first=first,
  decades=as.integer(seq(5, 45, 10)))
init.ricki5 <- function() list(lambda=runif(1, 10, 60), alpha=runif(1, 0, 60), 
  r=runif(1), K=runif(1, 30, 120), iota=runif(1),
  p0=rnorm(1, -1.5), p1=rnorm(1), p2=rnorm(1), p3=rnorm(1), p1st=rnorm(1), sigma.p=runif(1), N=N)
pars.ricki5 <- c("lambda", "alpha", "r", "K", "p0", "p1", "p2", "p3", "p1st", 
  "sigma.p", "iota", "Nmean", "Ndecade")

jm.ricki5 <- jags.model("dm.ricki5.txt", dat.ricki5, init.ricki5, n.chains=1, n.adapt=1000)
update(jm.ricki5, n.iter=100000)
jc.ricki5 <- coda.samples(jm.ricki5, pars.ricki5, n.iter=40000)

# Summarize results
mat.ricki5 <- as.matrix(jc.ricki5)
NmeanCols<- which(colnames(mat.ricki5)=="Nmean[1]") + 1:nYears - 1
Nmean.frame<-data.frame(year=1966:2010, Mean=colMeans(mat.ricki5[,NmeanCols]), HPDinterval(jc.ricki5[,NmeanCols]))
Nmean.frame
mean.N1 <- colMeans(mat.ricki5[,2:(122*5+1)])
mean.N1
mean.N2 <- matrix(mean.N1, nrow=122, ncol=5)
mean.N2
colnames(mean.N2) <- paste("N", seq(1970,2010,10), sep="")
max(mean.N2)
max.plot <- round(max(mean.N2)/5, -1)*5
mean.N2 <- with(rt2a, data.frame(statenum, Route, Lati, Longi, mean.N2))
head(mean.N2)

# Measures of variability
sd.N1 <- apply(mat.ricki5[,2:(122*45+1)], 2, sd)
sd.N1
mean(sd.N1)
mean(mean.N1)
mean(mat.ricki5[,"lambda"])
sd(mat.ricki5[,"lambda"])
sd(mat.ricki5[,"lambda"])/mean(mat.ricki5[,"lambda"])
sd(mat.ricki5[,"alpha"])/mean(mat.ricki5[,"alpha"])
sd(mat.ricki5[,"iota"])/mean(mat.ricki5[,"iota"])
sd(mat.ricki5[,"r"])/mean(mat.ricki5[,"r"])
summary(sd.N1/mean.N1)
sd.N2 <- apply(mean.N2, 2, sd) #Variation between sites in a year
sd.N2

# Make plot
library(ggplot2)
library(gridExtra)
library(maps)

all.states <- map_data("state")
states <- subset(all.states, region %in% c('maryland', 'virginia'))

max.radius <- 4
leg.sizes <- c(1, 10, 100, 250)
gN.sp <- list()
gN.sp[[1]] <- ggplot(mean.N2, aes(x=Longi,y=Lati)) + theme_bw() + labs(x="", y="")+
  geom_polygon(data=states, aes(x=long, y=lat, group=group), color="black", fill="white") +
  geom_point(aes(size=N1970), alpha=0.5, color="black")+scale_area(name="Abundance", range=c(1, max.radius), breaks=leg.sizes)+
  geom_text(x=-83.2, y=39.75, label="1970")
gN.sp[[2]] <- ggplot(mean.N2, aes(x=Longi,y=Lati)) + theme_bw() + labs(x="", y="")+
  geom_polygon(data=states, aes(x=long, y=lat, group=group), color="black", fill="white") +
  geom_point(aes(size=N1980), alpha=0.5, color="black")+scale_area(name="Abundance", range=c(1, max.radius), breaks=leg.sizes)+
  geom_text(x=-83.2, y=39.75, label="1980")
gN.sp[[3]] <- ggplot(mean.N2, aes(x=Longi,y=Lati)) + theme_bw() + labs(x="", y="")+
  geom_polygon(data=states, aes(x=long, y=lat, group=group), color="black", fill="white") +
  geom_point(aes(size=N1990), alpha=0.5, color="black")+scale_area(name="Abundance", range=c(1, max.radius), breaks=leg.sizes)+
  geom_text(x=-83.2, y=39.75, label="1990")
gN.sp[[4]] <- ggplot(mean.N2, aes(x=Longi,y=Lati)) + theme_bw() + labs(x="", y="")+
  geom_polygon(data=states, aes(x=long, y=lat, group=group), color="black", fill="white") +
  geom_point(aes(size=N2000), alpha=0.5, color="black")+scale_area(name="Abundance", range=c(1, max.radius), breaks=leg.sizes)+
  geom_text(x=-83.2, y=39.75, label="2000")
gN.sp[[5]] <- ggplot(mean.N2, aes(x=Longi,y=Lati)) + theme_bw() + labs(x="", y="")+
  geom_polygon(data=states, aes(x=long, y=lat, group=group), color="black", fill="white") +
  geom_point(aes(size=N2010), alpha=0.5, color="black")+scale_area(name="Abundance", range=c(1, max.radius), breaks=leg.sizes)+
  geom_text(x=-83.2, y=39.75, label="2010")

gN.mean <- ggplot(Nmean.frame.f, aes(x=year,y=Mean)) + theme_bw() + labs(x="Year",
  y="Mean Abundance")+
  geom_ribbon(aes(ymin=lower,ymax=upper),fill="black",
  alpha=0.5,colour=NA) + geom_line(colour="black", size=1)

png(filename = "OVEN_N_by_route_year6.png", height = 6.5, width = 6.5, units = "in", res=600)
print(arrangeGrob(gN.sp[[1]] + theme(legend.position=c(0.145, 0.56), 
  legend.background=element_blank(), axis.text=element_blank(), 
  axis.line=element_blank(), axis.ticks=element_blank(), panel.border=element_blank(),
  panel.grid=element_blank(), plot.margin=unit(rep(0.1, 4), "lines")),
  gN.sp[[2]] + theme(legend.position="none", axis.text=element_blank(), 
  axis.line=element_blank(), axis.ticks=element_blank(), panel.border=element_blank(),
  panel.grid=element_blank(), plot.margin=unit(rep(0.1, 4), "lines")),
  gN.sp[[3]] + theme(legend.position="none", axis.text=element_blank(), 
  axis.line=element_blank(), axis.ticks=element_blank(), panel.border=element_blank(),
  panel.grid=element_blank(), plot.margin=unit(rep(0.1, 4), "lines")),
  gN.sp[[4]] + theme(legend.position="none", axis.text=element_blank(), 
  axis.line=element_blank(), axis.ticks=element_blank(), panel.border=element_blank(),
  panel.grid=element_blank(), plot.margin=unit(rep(0.1, 4), "lines")),
  gN.sp[[5]] + theme(legend.position="none", axis.text=element_blank(), 
  axis.line=element_blank(), axis.ticks=element_blank(), panel.border=element_blank(),
  panel.grid=element_blank(), plot.margin=unit(rep(0.1, 4), "lines")),
  gN.mean, main=" ", nrow=3, ncol=2))
dev.off()

save.image(file.name)
\end{verbatim}

\end{document}